---
title: "Data Processing"
output: 
  html_document:
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
All data processing happens here: Products are saved in Processed data folder. All drought response data is based on BAI growth and lower 10th percentile spei years, but this can be changed here.

### Products
* detrended tree rings
* tree ring data joined with field data (diameter, competition etc.)


```{r packages, message=FALSE, warning=FALSE}
library(dplR)
library(tidyverse)
library(kableExtra)
library(knitr)
library(SPEI)
library(googledrive)
library(googlesheets4)
library(geosphere)
library(Hmisc)
library(reshape2)

```
## Site Characteristics
For consistent naming across data files. Note I'm also reading in the field data here to be manipulated later to keep the google code together

```{r google, echo=T, results='hide'}
## do authentication step first and copy id from google sheet url.
drive_auth(cache = ".secrets")
options(
  gargle_oauth_cache = ".secrets",
  gargle_oauth_email = "tbohn001@ucr.edu"
)

fielddata <- drive_get(as_id("1Y_PSgFJz4bSW00JYoRiVnXrCrhL776QA4dNSOWl4Clg"))
dia_data <- read_sheet(fielddata, sheet = 2)
sitechar <- read_sheet(fielddata, sheet=3)

sitechar <- rename(sitechar, siteno=No.) %>% 
  dplyr::select(c("Region", "siteno", "Site", "Site2", "Neighborhood", "hilo", "X", "Y"))

sites <- dplyr::select(sitechar, Site2, siteno, Region, hilo) %>% 
  group_by(Site2) %>% 
  summarise_all(first)

sites2 <- dplyr::select(sitechar, Site, Site2, siteno, Region, hilo) %>% 
  group_by(Site2) %>% 
  summarise_all(first)

```

## Tree Ring Data

#### Note on missing rings

When there is a missing ring in one of the two cores growth information is only taken from the core with growth information (not averaged).  When the ring is missing from both core samples the missing growth value is replaced with the mean growth for that tree but denoted as a missing ring for future analyses.

```{r data, echo=T, results='hide'}
bm <- read.rwl("Raw Data/RWL copies/Black_Mountain.rwl")
names(bm)[str_detect(names(bm), "11PL1_3A")==T] <- "11PL1_1C"
names(bm)[str_detect(names(bm), "11PL1_3B")==T] <- "11PL1_1D"

sp <- read.rwl("Raw Data/RWL copies/Seven_Pines.rwl")
bm_ids <- read.ids(bm, stc=c(2,5,1))
sp_ids <- read.ids(sp, stc=c(2,5,1))

bm_id_df <- bm_ids %>%
  mutate(label=rownames(bm_ids),
         Site="bm",
         Species=str_sub(label, 3, 4),
         Neighborhood=str_sub(label, 5, 5),
         ID=ifelse(str_detect(label, "\\_")==TRUE, str_sub(label, 7, 7), str_sub(label, 6,7)),
         tree=as.character(tree))

bm_id_df_tree <- group_by(bm_id_df, tree, Site, Species, Neighborhood, ID) %>%
  summarise(label=first(label))

sp_id_df <- sp_ids %>%
  mutate(label=rownames(sp_ids),
         Site="sp",
         Species=str_sub(label, 3, 4),
         Neighborhood=str_sub(label, 5, 5),
         ID=ifelse(str_detect(label, "\\_")==TRUE, str_sub(label, 7, 7), str_sub(label, 6,7)),
         tree=as.character(tree))
sp_id_df_tree <- group_by(sp_id_df, tree, Site, Species, Neighborhood, ID) %>%
  summarise(label=first(label))

lc <- read.rwl("Raw Data/RWL copies/Lilly_Creek.rwl")
pp <- read.rwl("Raw Data/RWL copies/Peppermint.rwl")
lc_ids <- read.ids(lc, stc=c(2,5,1))
pp_ids <- read.ids(pp, stc=c(2,5,1))

lc_id_df <- lc_ids %>%
  mutate(label=rownames(lc_ids),
         Site="lc",
         Species=str_sub(label, 3, 4),
         Neighborhood=str_sub(label, 5, 5), 
         ID=str_sub(label, 7, 7),
         tree=as.character(tree))
lc_id_df_tree <- group_by(lc_id_df, tree, Site, Species, Neighborhood, ID) %>%
  summarise(label=first(label))

pp_id_df <- pp_ids %>%
  mutate(label=rownames(pp_ids),
         Site="pp",
         Species=str_sub(label, 3, 4),
         Neighborhood=str_sub(label, 5, 5), 
         ID=str_sub(label, 7, 7),
         tree=as.character(tree))
pp_id_df_tree <- group_by(pp_id_df, tree, Site, Species, Neighborhood, ID) %>%
  summarise(label=first(label))

cm <- read.rwl("Raw Data/RWL copies/Crescent_Meadow.rwl")
names(cm)[str_detect(names(cm), "4PJ2_1A")==T] <- "4PP2_1A"
names(cm)[str_detect(names(cm), "4PJ2_1B")==T] <- "4PP2_1B"
pr <- read.rwl("Raw Data/RWL copies/Pine_Ridge.rwl")
cm_ids1 <- read.ids(cm[,which(str_length(colnames(cm))==7)], stc=c(1,5,1))
cm_ids2 <- read.ids(cm[,which(str_length(colnames(cm))==8)], stc=c(1,6,1))
cm_ids2$tree <- cm_ids2$tree + max(cm_ids1$tree)
cm_ids <- rbind(cm_ids1, cm_ids2)

pr_ids1 <- read.ids(pr[,which(str_length(colnames(pr))==7)], stc=c(1,5,1))
pr_ids2 <- read.ids(pr[,which(str_length(colnames(pr))==8)], stc=c(1,6,1))
pr_ids2$tree <- pr_ids2$tree + max(pr_ids1$tree)
pr_ids3 <- read.ids(pr[,which(str_length(colnames(pr))==6)], stc=c(1,5,5))
pr_ids3$tree <- pr_ids3$tree + max(pr_ids2$tree)
pr_ids <- rbind(pr_ids1, pr_ids2, pr_ids3)

cm_id_df <- cm_ids %>%
  mutate(label=rownames(cm_ids),
         Site="cm",
         Species=str_sub(label, 2, 3),
         tree=as.character(tree))
cm_id_df$Neighborhood <- NA
for(i in 1:nrow(cm_id_df)) {
  if(str_detect(cm_id_df$label[i], "_")==TRUE) {
    cm_id_df$Neighborhood[i] <- str_sub(cm_id_df$label[i], 4, 4)
  } else NA
}
cm_id_df$ID <- NA
for(i in 1:nrow(cm_id_df)) {
  if(str_detect(cm_id_df$label[i], "_")==TRUE) {
    cm_id_df$ID[i] <- str_sub(cm_id_df$label[i], 6, 6)
  } else cm_id_df$ID[i] <- str_sub(cm_id_df$label[i], 4, 7)
}
cm_id_df$ID <- str_replace_all(cm_id_df$ID, "A", "")
cm_id_df$ID <- str_replace_all(cm_id_df$ID, "B", "")
cm_id_df$Neighborhood <- ifelse(cm_id_df$ID%in%c(102, 105, 117, 118, 122), 1, 
                                ifelse(cm_id_df$ID%in%c(1924, 3064, 3085, 3101), 2,
                                       ifelse(cm_id_df$ID%in%c(1954, 1959, 1965, 1966, 1967), 3, cm_id_df$Neighborhood))) 

cm_id_df_tree <- group_by(cm_id_df, tree, Site, Species, Neighborhood, ID) %>%
  summarise(label=first(label))

pr_id_df <- pr_ids %>%
  mutate(label=rownames(pr_ids),
         Site="pr",
         Species=str_sub(label, 2, 3),
         tree=as.character(tree))
pr_id_df$Neighborhood <- NA
for(i in 1:nrow(pr_id_df)) {
  if(str_detect(pr_id_df$label[i], "_")==TRUE) {
    pr_id_df$Neighborhood[i] <- str_sub(pr_id_df$label[i], 4, 4)
  } else NA
}
pr_id_df$ID <- NA
for(i in 1:nrow(pr_id_df)) {
  if(str_detect(pr_id_df$label[i], "_")==TRUE) {
    pr_id_df$ID[i] <- str_sub(pr_id_df$label[i], 6, 6)
  } else pr_id_df$ID[i] <- str_sub(pr_id_df$label[i], 4, 7)
}
pr_id_df$ID <- str_replace_all(pr_id_df$ID, "A", "")
pr_id_df$ID <- str_replace_all(pr_id_df$ID, "B", "")

pr_id_df$ID[which(pr_id_df$ID==5434)] <- 543

pr_id_df$Neighborhood <- ifelse(pr_id_df$ID%in%c(497, 3150, 3157), 1,
                                ifelse(pr_id_df$ID%in%c(528, 533, 543, 552, 3001), 2,
                                       ifelse(pr_id_df$ID%in%c(598, 604), 3, pr_id_df$Neighborhood)))

pr_id_df_tree <- group_by(pr_id_df, tree, Site, Species, Neighborhood, ID) %>%
  summarise(label=first(label))


ic <- read.rwl("Raw Data/RWL copies/Inyo_Craters.rwl")
names(ic)[str_detect(names(ic), "9AC1_1A")==T] <- "9AM1_1A"
names(ic)[str_detect(names(ic), "9AC1_1B")==T] <- "9AM1_1B"
sl <- read.rwl("Raw Data/RWL copies/Sherwin_Lakes.rwl")
ic_ids <- read.ids(ic, stc=c(1,5,1))
sl_ids <- read.ids(sl, stc=c(2,5,1))

ic_id_df <- ic_ids %>%
  mutate(label=rownames(ic_ids),
         Site="ic",
         Species=str_sub(label, 2, 3),
         Neighborhood=str_sub(label, 4, 4), 
         ID=str_sub(label, 6, 6),
         tree=as.character(tree))
ic_id_df_tree <- group_by(ic_id_df, tree, Site, Species, Neighborhood, ID) %>%
  summarise(label=first(label))

sl_id_df <- sl_ids %>%
  mutate(label=rownames(sl_ids),
         Site="sl",
         Species=str_sub(label, 3, 4),
         Neighborhood=str_sub(label, 5, 5), 
         ID=str_sub(label, 7, 7),
         tree=as.character(tree))
sl_id_df_tree <- group_by(sl_id_df, tree, Site, Species, Neighborhood, ID) %>%
  summarise(label=first(label))

for(i in c("sp", "bm", "lc", "pp", "pr", "cm", "ic", "sl")){ 
  temp <- get(i)
  temp[temp == 0] <- NA
  treeave <- treeMean(temp, get(paste(i,"_ids", sep="")), na.rm = T)
  zero <- fill.internal.NA(treeave, 0)
  zero$Year <- as.numeric(rownames(zero))

  treeave <- fill.internal.NA(treeave, "Mean")
  raw <- treeave
  raw$Year <- as.numeric(rownames(raw))
  
  interseries.cor(temp, prewhiten = T)
  corr.rwl.seg(temp,prewhiten = T)
  rwi.stats.running(raw, prewhiten=T)
  
  bai <- bai.out(raw, diam = NULL)
  bai$Year <- as.numeric(rownames(bai))
  rwi.stats(bai, prewhiten=T)
  
  ar <- detrend(rwl = treeave, method="Ar")
  ar$Year <- as.numeric(rownames(ar))
  
  nexp <- detrend(rwl = treeave, method="ModNegExp")
  nexp$Year <- as.numeric(rownames(nexp))
  
  rwi.stats(nexp, prewhiten=T)
  
  spl <- detrend(rwl = treeave, method="Spline")
  spl$Year <- as.numeric(rownames(spl))
  
  rwi.stats(spl, prewhiten=T)
  
  fss <- detrend(rwl = treeave, method="Friedman")
  fss_ar <- detrend(fss, method="Ar")
  fss$Year <- as.numeric(rownames(fss))
  
  rwi.stats(fss, prewhiten=T)
  
  dat<- gather(zero, key="tree", "temp", -Year) %>%
    mutate(zeroRing=ifelse(temp==0, "yes", "no")) %>% 
    dplyr::select(-temp) %>% 
    left_join(gather(raw, key="tree", "raw_growth", -Year)) %>%
    left_join(gather(bai, key="tree", "BAI", -Year)) %>%
    left_join(gather(nexp, key="tree", "nexp_growth", -Year)) %>%
    left_join(gather(spl, key="tree", "spline_growth", -Year)) %>%
    left_join(gather(fss, key="tree", "fss_growth", -Year)) %>%
    na.omit() %>%
    left_join(get(paste(i,"_id_df_tree", sep=""))) %>%
    filter(Year>=1895)
  
  if(i=="sp"){allrings_loop <- dat
  } else allrings_loop <- rbind(allrings_loop, dat)
  
}

allrings <- allrings_loop %>% 
  left_join(sites, by=c("Site"="Site2")) %>% 
  mutate(tree.uniqueID = tolower(paste(paste(siteno, Species, Neighborhood, sep=""), ID, sep="_"))) %>% 
  dplyr::select( -tree) %>% 
  mutate_at(c("Neighborhood", "ID"), as.numeric)

write.csv(allrings, "Processed Data/detrended_rings_allsites.csv", row.names=F)
```

## Crossdating summary stats
```{r summarystats, echo=FALSE}
sj_summary <- bind_rows(rwi.stats(sp, sp_ids, prewhiten=T)%>% mutate(site='seven pines'),
                        rwi.stats(bm, bm_ids, prewhiten=T)%>% mutate(site='black mountain'))
sj_summary <- sj_summary[c(14, 1:13)]
kable(sj_summary, caption = "San Jacinto") %>% kable_styling()

senf_summary <- bind_rows(rwi.stats(lc, lc_ids, prewhiten=T)%>% mutate(site='lilly creek'),
                        rwi.stats(pp, pp_ids, prewhiten=T)%>% mutate(site='peppermint'))
senf_summary <- senf_summary[c(14, 1:13)]
kable(senf_summary, caption = "Sequoia National Forest") %>% kable_styling()

seki_summary <- bind_rows(rwi.stats(cm, cm_ids, prewhiten=T)%>% mutate(site='crescent meadow'),
                        rwi.stats(pr, pr_ids, prewhiten=T)%>% mutate(site='pine ridge'))
seki_summary <- seki_summary[c(14, 1:13)]
kable(seki_summary, caption = "Sequoia National Park") %>% kable_styling()

mamm_summary <- bind_rows(rwi.stats(ic, ic_ids, prewhiten=T)%>% mutate(site='inyo craters'),
                        rwi.stats(sl, sl_ids, prewhiten=T)%>% mutate(site='sherwin lakes'))
mamm_summary <- mamm_summary[c(14, 1:13)]
kable(mamm_summary, caption = "Mammoth Lakes") %>% kable_styling()
```


## Field data
This includes location, DBH, site characteristics, and hegyi index calculation.  Still a handful of id problems, verify field and gps data.

```{r manip, echo=T, results='hide'}
dia_data2 <- dia_data %>% 
  dplyr::select(-Region) %>% 
  left_join(sites2) %>% 
  mutate(Species=tolower(paste(str_sub(Species, 1,1), str_sub(Species, 3,3), sep="")),
         Species=ifelse(Species=="?", "xx",
                         ifelse(Species=="x", "xx",
                                ifelse(Species=="a?", "ax", Species))),
         tree.uniqueID = tolower(paste(paste(siteno, Species, Neighborhood, sep=""), ID, sep="_")),
         DBHstem1=ifelse(!is.na(DBHT3), DBHT3, 
                       ifelse(!is.na(DBHT2), DBHT2, DBH)),
         BA1 = .00007854*DBHstem1^2,
         BA2 = .00007854*DBHstem2^2,
         BA3 = .00007854*DBHstem3^2) %>% 
  mutate(totDBH=rowSums(dplyr::select(.,"DBHstem1", "DBHstem2", "DBHstem3"), na.rm=T),
         totBA=rowSums(dplyr::select(.,"BA1", "BA2", "BA3"), na.rm=T),
         totDBH=ifelse(totDBH==0, NA, totDBH),
         totBA=ifelse(totBA==0, NA, totBA)) %>% 
  dplyr::select(c(tree.uniqueID, Site, Neighborhood, Species, totDBH, totBA))


raw_loc <- read.csv("Raw data/all_loc.csv") 

test <- raw_loc %>% 
  filter(tree.uniqueID %in% raw_loc$tree.uniqueID[duplicated(raw_loc$tree.uniqueID)==T])

loc_data <- raw_loc %>% 
  inner_join(dplyr::select(dia_data2, c(tree.uniqueID, Site, Neighborhood))) %>% ## there are some naming errors
  group_by(tree.uniqueID) %>% summarise_all(first)
# loc_data <- loc_data %>%
#   filter(Site=="Black Mountain") %>%
#   filter(Neighborhood==3)

loc_grid <- loc_data %>% 
  rename(Site.x=Site, Neighborhood.x=Neighborhood, tree.uniqueID.x=tree.uniqueID, X.x=X, Y.x=Y) %>% 
  expand_grid(dplyr::select(loc_data, c(Site, Neighborhood, tree.uniqueID, X, Y))) %>% 
  filter(Site.x==Site, Neighborhood.x==Neighborhood) %>% 
  dplyr::select(-c(Site.x, Neighborhood.x)) 

loc_grid2 <- loc_grid %>% 
  mutate(dist_m=distGeo(dplyr::select(loc_grid, c(X.x, Y.x)), dplyr::select(loc_grid, c(X, Y))),
         bearing_deg=round(bearing(dplyr::select(loc_grid, c(X.x, Y.x)), dplyr::select(loc_grid, c(X, Y))), 2))

xy <- dplyr::select(loc_data, c(X, Y))
pair <- data.frame(distm(xy, fun=distGeo)) %>% ## calculates distance in m based on WGS84 elipsoid
  mutate(tree.uniqueID=loc_data$tree.uniqueID)
names(pair) <- c(loc_data$tree.uniqueID, "focalID")

test <- xy %>% 
  mutate(bearing_deg=round(bearing(xy), 2)) %>% 
  left_join(loc_data)

dist_focal <- pair %>% 
  pivot_longer(cols=-focalID, names_to="periphID", values_to="dist") %>% 
  filter(dist!=0) %>% 
  filter(focalID %in% unique(allrings$tree.uniqueID)) %>% 
  filter(periphID %in% unique(allrings$tree.uniqueID)) %>% 
  left_join(dplyr::select(dia_data2, c("focalID"="tree.uniqueID", Species, totDBH, Site, Neighborhood))) %>% 
  rename(focalsp=Species, DBH=totDBH, Site.x=Site, Neighborhood.x=Neighborhood) %>% 
  left_join(dplyr::select(dia_data2, c("periphID"="tree.uniqueID", Species, totBA, Site, Neighborhood))) %>% 
  rename(periphsp=Species) %>% 
  filter(Site==Site.x, Neighborhood==Neighborhood.x) %>% 
  dplyr::select(-c(Neighborhood.x, Site.x, Neighborhood, Site))

dist_long <- pair %>% 
  pivot_longer(cols=-focalID, names_to="periphID", values_to="dist") %>% 
  filter(dist!=0, dist<=20) %>% 
  filter(focalID %in% unique(allrings$tree.uniqueID)) %>% 
  left_join(dplyr::select(dia_data2, c("focalID"="tree.uniqueID", Species, totDBH, totBA))) %>% 
  rename(focalsp=Species, DBH=totDBH, focalBA=totBA) %>% 
  left_join(dplyr::select(dia_data2, c("periphID"="tree.uniqueID", Species, totBA))) %>% 
  rename(periphsp=Species) %>% 
  mutate(con_het=ifelse(focalsp==periphsp, "con", "het"))

comp_summary <- dist_long %>% 
  group_by(focalID, DBH, focalBA, con_het) %>% 
  summarise(stems=length(periphID),
            BA=sum(totBA)) %>% 
  pivot_wider(id_cols = c(focalID, DBH, focalBA), names_from = con_het, values_from=c(stems, BA)) %>% 
  mutate_at(vars(-group_cols()), .funs=function(x){ifelse(is.na(x), 0, x)})

freq_dat <- comp_summary %>% 
  left_join(dplyr::select(dia_data2, c("focalID"="tree.uniqueID", Site, Neighborhood))) %>% 
  group_by(focalID) %>% 
  mutate(con_freq=(1+stems_con)/(1+stems_con+stems_het),
         con_freqBA=(focalBA+BA_con)/(focalBA+BA_con+BA_het))

write.csv(dist_focal, "Processed Data/dist_focal.csv", row.names=F)
write.csv(dist_long, "Processed Data/dist_long.csv", row.names=F)
write.csv(freq_dat, "Processed Data/frequency_data.csv", row.names = F)

rings_field <- left_join(allrings, dplyr::select(freq_dat, -c(Site, Neighborhood)), by=c("tree.uniqueID"="focalID"))

write.csv(rings_field, "Processed Data/matched_rings_field.csv", row.names = F)
```

#### Synchrony 
Distribution of pairwise pearson correlations between tree ring series broken down by species and site (based on spline detrended series).
```{r synchrony, echo=F}
rings_field <- rings_field %>% 
  group_by(tree.uniqueID) %>% 
  mutate(minyear=min(Year)) %>% 
  ungroup() %>% as.data.frame()

for(i in unique(rings_field$Region)){
  tree_sub <- filter(rings_field, Region==i) %>% 
      # filter(Year>=max(minyear)) %>%
      dplyr::select(tree.uniqueID, Year, spline_growth) %>% 
      pivot_wider(names_from="tree.uniqueID", values_from="spline_growth") 
    
    if(dim(tree_sub)[1]==0){    
      cordf <- NA
    } else {
        treemat <-tree_sub %>% 
          dplyr::select(-Year) %>%  as.matrix()

        cormat <- rcorr(treemat, type='pearson')$r
        cormat[upper.tri(cormat)] <- NA

       cordf <- melt(cormat) %>% filter(value!=1&!is.na(value)) %>% 
          rename(tree1=Var1, tree2=Var2, pearson_r=value) 
      }

    
    if(i=="San Jac"){all_cordf <- cordf
      } else all_cordf <- rbind(all_cordf, cordf)
}

sub <- dplyr::select(rings_field, c(tree.uniqueID, Site, Species, Region, hilo, Neighborhood, DBH, con_freq, con_freqBA)) %>% 
  group_by(tree.uniqueID) %>% 
  summarise_all(first)

dist_mat2 <- dplyr::select(dist_focal, c(focalID, periphID, dist)) %>% 
  rename(tree1=focalID, tree2=periphID)

all_cordf2 <- all_cordf %>% 
  na.omit() %>% 
  left_join(sub, by=c("tree1"="tree.uniqueID")) %>% 
  left_join(dplyr::select(sub, c(tree.uniqueID, Site, Neighborhood, Species)), by=c("tree2"="tree.uniqueID")) %>% 
  left_join(dist_mat2) %>% 
  filter(Neighborhood.x==Neighborhood.y, Site.x==Site.y)


all_cordf3 <- all_cordf2 %>% 
  na.omit()


write.csv(all_cordf2, "Processed Data/synchrony_dat.csv", row.names = F)


for(i in unique(rings_field$Region)){
  tree_sub <- filter(rings_field, Region==i) %>% 
      dplyr::select(tree.uniqueID, Year, spline_growth) %>% 
      pivot_wider(names_from="tree.uniqueID", values_from="spline_growth") 
  
  for(y in seq(1910, 2020, by=10)){
    tree_sub2 <- tree_sub %>% 
      filter(Year<=y, Year>=(y-29))
    if(dim(tree_sub2)[1]<20){    
      cordf <- data.frame(tree1=NA, tree2=NA, pearson_r=NA, decade=y)
    } else {
        treemat <-tree_sub2 %>% 
          dplyr::select(-Year) %>%  as.matrix()

        cormat <- rcorr(treemat, type='pearson')$r
        cormat[upper.tri(cormat)] <- NA

       cordf <- melt(cormat) %>% filter(value!=1&!is.na(value)) %>% 
         rename(tree1=Var1, tree2=Var2, pearson_r=value) %>% 
         mutate(decade=y)
      }

    
    if(i=="San Jac"& y==1910){all_cordf <- cordf
      } else all_cordf <- rbind(all_cordf, cordf)
}}

sub <- dplyr::select(rings_field, c(tree.uniqueID, Site, Species, Region, hilo, Neighborhood, DBH, con_freq, con_freqBA)) %>% 
  group_by(tree.uniqueID) %>% 
  summarise_all(first)

dist_mat2 <- dplyr::select(dist_focal, c(focalID, periphID, dist)) %>% 
  rename(tree1=focalID, tree2=periphID)

all_cordf2 <- all_cordf %>% 
  na.omit() %>% 
  left_join(sub, by=c("tree1"="tree.uniqueID")) %>% 
  left_join(dplyr::select(sub, c(tree.uniqueID, Site, Neighborhood, Species)), by=c("tree2"="tree.uniqueID")) %>% 
  left_join(dist_mat2) %>% 
  filter(Neighborhood.x==Neighborhood.y, Site.x==Site.y)


all_cordf3 <- all_cordf2 %>% 
  na.omit()


write.csv(all_cordf2, "Processed Data/synchrony_decadal_pearson.csv", row.names = F)

for(i in unique(rings_field$Region)){
  tree_sub <- filter(rings_field, Region==i) %>% 
      filter(Year>=max(minyear)) %>%
      dplyr::select(tree.uniqueID, Year, raw_growth) %>% 
      pivot_wider(names_from="tree.uniqueID", values_from="raw_growth") 
    
    if(dim(tree_sub)[1]==0){    
      cordf <- NA
    } else {
        treemat <-tree_sub %>% 
          dplyr::select(-Year) %>%  as.matrix()

        cormat <- rcorr(treemat, type='pearson')$r
        cormat[upper.tri(cormat)] <- NA

       cordf <- melt(cormat) %>% filter(value!=1&!is.na(value)) %>% 
          rename(tree1=Var1, tree2=Var2, pearson_r=value) 
      }

    
    if(i=="San Jac"){all_cordf <- cordf
      } else all_cordf <- rbind(all_cordf, cordf)
}


all_cordf_raw <- all_cordf %>% 
  na.omit() %>% 
  mutate_at("tree1", as.character) %>% 
  left_join(sub, by=c("tree1"="tree.uniqueID")) %>% 
  left_join(dplyr::select(sub, c(tree.uniqueID, Site, Neighborhood, Species)), by=c("tree2"="tree.uniqueID")) %>% 
  left_join(dist_mat2) %>% 
  filter(Neighborhood.x==Neighborhood.y, Site.x==Site.y)

write.csv(all_cordf_raw, "Processed Data/synchrony_raw.csv", row.names = F)

```

