---
title: "Data Processing"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
All data processing happens here: Products are saved in Processed data folder. All drought response data is based on BAI growth and lower 10th percentile spei years, but this can be changed here.

### Products
* detrended tree rings
* tree ring data joined with field data (diameter, competition etc.)


```{r packages, message=FALSE, warning=FALSE}
library(dplR)
library(tidyverse)
library(kableExtra)
library(knitr)
library(SPEI)
library(googledrive)
library(googlesheets4)

```
## Site Characteristics
For consistent naming across data files. Note I'm also reading in the field data here to be manipulated later to keep the google code together

```{r google, echo=T, results='hide'}
## do authentication step first and copy id from google sheet url.
drive_auth(cache = ".secrets")
options(
  gargle_oauth_cache = ".secrets",
  gargle_oauth_email = "tbohn001@ucr.edu"
)

fielddata <- drive_get(as_id("1Y_PSgFJz4bSW00JYoRiVnXrCrhL776QA4dNSOWl4Clg"))
dia_data <- read_sheet(fielddata, sheet = 2)
sitechar <- read_sheet(fielddata, sheet=3)

sitechar <- rename(sitechar, siteno=No.) %>% 
  dplyr::select(c("Region", "siteno", "Site", "Site2", "Neighborhood", "hilo", "X", "Y"))

sites <- dplyr::select(sitechar, Site2, siteno, Region, hilo) %>% 
  group_by(Site2) %>% 
  summarise_all(first)

sites2 <- dplyr::select(sitechar, Site, Site2, siteno, Region, hilo) %>% 
  group_by(Site2) %>% 
  summarise_all(first)

```

## Tree Ring Data

#### Note on missing rings

When there is a missing ring in one of the two cores growth information is only taken from the core with growth information (not averaged).  When the ring is missing from both core samples the missing growth value is replaced with the mean growth for that tree but denoted as a missing ring for future analyses.

```{r data, echo=T, results='hide'}
bm <- read.rwl("Raw Data/RWL copies/Black_Mountain.rwl")
names(bm)[str_detect(names(bm), "11PL1_3A")==T] <- "11PL1_1C"
names(bm)[str_detect(names(bm), "11PL1_3B")==T] <- "11PL1_1D"

sp <- read.rwl("Raw Data/RWL copies/Seven_Pines.rwl")
bm_ids <- read.ids(bm, stc=c(2,5,1))
sp_ids <- read.ids(sp, stc=c(2,5,1))

bm_id_df <- bm_ids %>%
  mutate(label=rownames(bm_ids),
         Site="bm",
         Species=str_sub(label, 3, 4),
         Neighborhood=str_sub(label, 5, 5),
         ID=ifelse(str_detect(label, "\\_")==TRUE, str_sub(label, 7, 7), str_sub(label, 6,7)),
         tree=as.character(tree))

bm_id_df_tree <- group_by(bm_id_df, tree, Site, Species, Neighborhood, ID) %>%
  summarise(label=first(label))

sp_id_df <- sp_ids %>%
  mutate(label=rownames(sp_ids),
         Site="sp",
         Species=str_sub(label, 3, 4),
         Neighborhood=str_sub(label, 5, 5),
         ID=ifelse(str_detect(label, "\\_")==TRUE, str_sub(label, 7, 7), str_sub(label, 6,7)),
         tree=as.character(tree))
sp_id_df_tree <- group_by(sp_id_df, tree, Site, Species, Neighborhood, ID) %>%
  summarise(label=first(label))

lc <- read.rwl("Raw Data/RWL copies/Lilly_Creek.rwl")
pp <- read.rwl("Raw Data/RWL copies/Peppermint.rwl")
lc_ids <- read.ids(lc, stc=c(2,5,1))
pp_ids <- read.ids(pp, stc=c(2,5,1))

lc_id_df <- lc_ids %>%
  mutate(label=rownames(lc_ids),
         Site="lc",
         Species=str_sub(label, 3, 4),
         Neighborhood=str_sub(label, 5, 5), 
         ID=str_sub(label, 7, 7),
         tree=as.character(tree))
lc_id_df_tree <- group_by(lc_id_df, tree, Site, Species, Neighborhood, ID) %>%
  summarise(label=first(label))

pp_id_df <- pp_ids %>%
  mutate(label=rownames(pp_ids),
         Site="pp",
         Species=str_sub(label, 3, 4),
         Neighborhood=str_sub(label, 5, 5), 
         ID=str_sub(label, 7, 7),
         tree=as.character(tree))
pp_id_df_tree <- group_by(pp_id_df, tree, Site, Species, Neighborhood, ID) %>%
  summarise(label=first(label))

cm <- read.rwl("Raw Data/RWL copies/Crescent_Meadow.rwl")
names(cm)[str_detect(names(cm), "4PJ2_1A")==T] <- "4PP2_1A"
names(cm)[str_detect(names(cm), "4PJ2_1B")==T] <- "4PP2_1B"
pr <- read.rwl("Raw Data/RWL copies/Pine_Ridge.rwl")
cm_ids1 <- read.ids(cm[,which(str_length(colnames(cm))==7)], stc=c(1,5,1))
cm_ids2 <- read.ids(cm[,which(str_length(colnames(cm))==8)], stc=c(1,6,1))
cm_ids2$tree <- cm_ids2$tree + max(cm_ids1$tree)
cm_ids <- rbind(cm_ids1, cm_ids2)

pr_ids1 <- read.ids(pr[,which(str_length(colnames(pr))==7)], stc=c(1,5,1))
pr_ids2 <- read.ids(pr[,which(str_length(colnames(pr))==8)], stc=c(1,6,1))
pr_ids2$tree <- pr_ids2$tree + max(pr_ids1$tree)
pr_ids3 <- read.ids(pr[,which(str_length(colnames(pr))==6)], stc=c(1,5,5))
pr_ids3$tree <- pr_ids3$tree + max(pr_ids2$tree)
pr_ids <- rbind(pr_ids1, pr_ids2, pr_ids3)

cm_id_df <- cm_ids %>%
  mutate(label=rownames(cm_ids),
         Site="cm",
         Species=str_sub(label, 2, 3),
         tree=as.character(tree))
cm_id_df$Neighborhood <- NA
for(i in 1:nrow(cm_id_df)) {
  if(str_detect(cm_id_df$label[i], "_")==TRUE) {
    cm_id_df$Neighborhood[i] <- str_sub(cm_id_df$label[i], 4, 4)
  } else NA
}
cm_id_df$ID <- NA
for(i in 1:nrow(cm_id_df)) {
  if(str_detect(cm_id_df$label[i], "_")==TRUE) {
    cm_id_df$ID[i] <- str_sub(cm_id_df$label[i], 6, 6)
  } else cm_id_df$ID[i] <- str_sub(cm_id_df$label[i], 4, 7)
}
cm_id_df$ID <- str_replace_all(cm_id_df$ID, "A", "")
cm_id_df$ID <- str_replace_all(cm_id_df$ID, "B", "")
cm_id_df$Neighborhood <- ifelse(cm_id_df$ID%in%c(102, 105, 117, 118, 122), 1, 
                                ifelse(cm_id_df$ID%in%c(1924, 3064, 3085, 3101), 2,
                                       ifelse(cm_id_df$ID%in%c(1954, 1959, 1965, 1966, 1967), 3, cm_id_df$Neighborhood))) 

cm_id_df_tree <- group_by(cm_id_df, tree, Site, Species, Neighborhood, ID) %>%
  summarise(label=first(label))

pr_id_df <- pr_ids %>%
  mutate(label=rownames(pr_ids),
         Site="pr",
         Species=str_sub(label, 2, 3),
         tree=as.character(tree))
pr_id_df$Neighborhood <- NA
for(i in 1:nrow(pr_id_df)) {
  if(str_detect(pr_id_df$label[i], "_")==TRUE) {
    pr_id_df$Neighborhood[i] <- str_sub(pr_id_df$label[i], 4, 4)
  } else NA
}
pr_id_df$ID <- NA
for(i in 1:nrow(pr_id_df)) {
  if(str_detect(pr_id_df$label[i], "_")==TRUE) {
    pr_id_df$ID[i] <- str_sub(pr_id_df$label[i], 6, 6)
  } else pr_id_df$ID[i] <- str_sub(pr_id_df$label[i], 4, 7)
}
pr_id_df$ID <- str_replace_all(pr_id_df$ID, "A", "")
pr_id_df$ID <- str_replace_all(pr_id_df$ID, "B", "")

pr_id_df$ID[which(pr_id_df$ID==5434)] <- 543

pr_id_df$Neighborhood <- ifelse(pr_id_df$ID%in%c(497, 3150, 3157), 1,
                                ifelse(pr_id_df$ID%in%c(528, 533, 543, 552, 3001), 2,
                                       ifelse(pr_id_df$ID%in%c(598, 604), 3, pr_id_df$Neighborhood)))

pr_id_df_tree <- group_by(pr_id_df, tree, Site, Species, Neighborhood, ID) %>%
  summarise(label=first(label))


ic <- read.rwl("Raw Data/RWL copies/Inyo_Craters.rwl")
names(ic)[str_detect(names(ic), "9AC1_1A")==T] <- "9AM1_1A"
names(ic)[str_detect(names(ic), "9AC1_1B")==T] <- "9AM1_1B"
sl <- read.rwl("Raw Data/RWL copies/Sherwin_Lakes.rwl")
ic_ids <- read.ids(ic, stc=c(1,5,1))
sl_ids <- read.ids(sl, stc=c(2,5,1))

ic_id_df <- ic_ids %>%
  mutate(label=rownames(ic_ids),
         Site="ic",
         Species=str_sub(label, 2, 3),
         Neighborhood=str_sub(label, 4, 4), 
         ID=str_sub(label, 6, 6),
         tree=as.character(tree))
ic_id_df_tree <- group_by(ic_id_df, tree, Site, Species, Neighborhood, ID) %>%
  summarise(label=first(label))

sl_id_df <- sl_ids %>%
  mutate(label=rownames(sl_ids),
         Site="sl",
         Species=str_sub(label, 3, 4),
         Neighborhood=str_sub(label, 5, 5), 
         ID=str_sub(label, 7, 7),
         tree=as.character(tree))
sl_id_df_tree <- group_by(sl_id_df, tree, Site, Species, Neighborhood, ID) %>%
  summarise(label=first(label))

for(i in c("sp", "bm", "lc", "pp", "pr", "cm", "ic", "sl")){ 
  temp <- get(i)
  temp[temp == 0] <- NA
  treeave <- treeMean(temp, get(paste(i,"_ids", sep="")), na.rm = T)
  zero <- fill.internal.NA(treeave, 0)
  zero$Year <- as.numeric(rownames(zero))

  treeave <- fill.internal.NA(treeave, "Mean")
  raw <- treeave
  raw$Year <- as.numeric(rownames(raw))
  
  bai <- bai.out(raw, diam = NULL)
  bai$Year <- as.numeric(rownames(bai))
  
  spl <- detrend(rwl = treeave, method="Spline")
  spl$Year <- as.numeric(rownames(spl))
  
  fss <- detrend(rwl = treeave, method="Friedman")
  fss$Year <- as.numeric(rownames(fss))
  
  dat<- gather(zero, key="tree", "temp", -Year) %>%
    mutate(zeroRing=ifelse(temp==0, "yes", "no")) %>% 
    dplyr::select(-temp) %>% 
    left_join(gather(raw, key="tree", "raw_growth", -Year)) %>%
    left_join(gather(bai, key="tree", "BAI", -Year)) %>%
    left_join(gather(spl, key="tree", "spline_growth", -Year)) %>%
    left_join(gather(fss, key="tree", "fss_growth", -Year)) %>%
    na.omit() %>%
    left_join(get(paste(i,"_id_df_tree", sep=""))) %>%
    filter(Year>=1895)
  
  if(i=="sp"){allrings_loop <- dat
  } else allrings_loop <- rbind(allrings_loop, dat)
  
}

allrings <- allrings_loop %>% 
  left_join(sites, by=c("Site"="Site2")) %>% 
  mutate(tree.uniqueID = tolower(paste(paste(siteno, Species, Neighborhood, sep=""), ID, sep="_"))) %>% 
  dplyr::select( -tree) %>% 
  mutate_at(c("Neighborhood", "ID"), as.numeric)

write.csv(allrings, "Processed Data/detrended_rings_allsites.csv", row.names=F)
```

## Crossdating summary stats
```{r summarystats, echo=FALSE}
sj_summary <- bind_rows(rwi.stats(sp, sp_ids, prewhiten=T)%>% mutate(site='seven pines'),
                        rwi.stats(bm, bm_ids, prewhiten=T)%>% mutate(site='black mountain'))
sj_summary <- sj_summary[c(14, 1:13)]
kable(sj_summary, caption = "San Jacinto") %>% kable_styling()

senf_summary <- bind_rows(rwi.stats(lc, lc_ids, prewhiten=T)%>% mutate(site='lilly creek'),
                        rwi.stats(pp, pp_ids, prewhiten=T)%>% mutate(site='peppermint'))
senf_summary <- senf_summary[c(14, 1:13)]
kable(senf_summary, caption = "Sequoia National Forest") %>% kable_styling()

seki_summary <- bind_rows(rwi.stats(cm, cm_ids, prewhiten=T)%>% mutate(site='crescent meadow'),
                        rwi.stats(pr, pr_ids, prewhiten=T)%>% mutate(site='pine ridge'))
seki_summary <- seki_summary[c(14, 1:13)]
kable(seki_summary, caption = "Sequoia National Park") %>% kable_styling()

mamm_summary <- bind_rows(rwi.stats(ic, ic_ids, prewhiten=T)%>% mutate(site='inyo craters'),
                        rwi.stats(sl, sl_ids, prewhiten=T)%>% mutate(site='sherwin lakes'))
mamm_summary <- mamm_summary[c(14, 1:13)]
kable(mamm_summary, caption = "Mammoth Lakes") %>% kable_styling()
```


## Field data
This includes location, DBH, site characteristics, and hegyi index calculation.  Still a handful of id problems, verify field and gps data.

```{r manip, echo=T, results='hide'}
dia_data2 <- dia_data %>% 
  dplyr::select(-Region) %>% 
  left_join(sites2) %>% 
  mutate(Species=tolower(paste(str_sub(Species, 1,1), str_sub(Species, 3,3), sep="")),
         Species=ifelse(Species=="?", "xx",
                         ifelse(Species=="x", "xx",
                                ifelse(Species=="a?", "ax", Species))),
         tree.uniqueID = tolower(paste(paste(siteno, Species, Neighborhood, sep=""), ID, sep="_")),
         DBHstem1=ifelse(!is.na(DBHT3), DBHT3, 
                       ifelse(!is.na(DBHT2), DBHT2, DBH)),
         BA1 = .00007854*DBHstem1^2,
         BA2 = .00007854*DBHstem2^2,
         BA3 = .00007854*DBHstem3^2) %>% 
  mutate(totDBH=rowSums(dplyr::select(.,"DBHstem1", "DBHstem2", "DBHstem3"), na.rm=T),
         totBA=rowSums(dplyr::select(.,"BA1", "BA2", "BA3"), na.rm=T),
         totDBH=ifelse(totDBH==0, NA, totDBH),
         totBA=ifelse(totBA==0, NA, totBA)) %>% 
  dplyr::select(c(tree.uniqueID, Site, Neighborhood, Species, totDBH, totBA))

freq_dat <- dia_data2 %>% 
  group_by(Site, Neighborhood) %>% 
  mutate(nb_BA=sum(totBA, na.rm=T),
         nb_stem=length(!is.na(totBA))) %>% 
  group_by(Site, Neighborhood, Species) %>% 
  summarise(nb_BA=mean(nb_BA),
            nb_stem=mean(nb_stem),
            sp_BA=sum(totBA, na.rm=T),
            sp_stem=length(!is.na(totBA)),
            freq_BA=sp_BA/nb_BA,
            freq_stem=sp_stem/nb_stem)

write.csv(freq_dat, "Processed Data/frequency_data.csv", row.names = F)
# 
# focalsub <- group_by(allrings, tree.uniqueID) %>%
#   summarise()
# 
# test <-anti_join(focalsub, dia_data2)
dia_data2 <- dplyr::select(dia_data2, c(tree.uniqueID, totDBH, totBA))
dist_mat <- read.csv("Raw Data/dist mat 20m.csv") %>% 
  rename(focalID=InputID, compID=TargetID) %>% 
  mutate_at(c("focalID", "compID"), as.character) %>% 
  left_join(dia_data2, by=c("focalID"="tree.uniqueID")) %>%
  rename(focalDBH=totDBH, focalBA=totBA) %>% 
  left_join(dia_data2, by=c("compID"="tree.uniqueID")) %>%
  rename(compDBH=totDBH, compBA=totBA) %>% 
  filter(focalID!=compID) %>% 
  mutate(Distance=ifelse(Distance==0, 0.000001, Distance))

# treeswithcoord <- group_by(dist_mat, focalID) %>%
#   summarise() %>%
#   filter(focalID%in%unique(allrings$tree.uniqueID))
# 
# test2 <- anti_join(focalsub, treeswithcoord, by=c("tree.uniqueID"="focalID"))

hegyi <- dist_mat %>%
  # na.omit() %>% 
  group_by(focalID, focalDBH) %>%
  mutate(ratio=(compDBH/focalDBH)/Distance,
         simple=compDBH/Distance,
         BA= compBA) %>%
  summarise(hegyi=sum(ratio, na.rm=T),
            simple_comp=(sum(simple, na.rm=T)),
            BA = sum(BA, na.rm=T)) ### missing lots of dbh's

hegyi_focal <- filter(hegyi, focalID%in%unique(allrings$tree.uniqueID)) %>%
  rename(tree.uniqueID=focalID) %>% 
  dplyr::select(-focalDBH)

rings_field <- inner_join(allrings, hegyi_focal) %>%
  left_join(dia_data2) %>%
  rename(DBH=totDBH)

write.csv(rings_field, "Processed Data/matched_rings_field.csv", row.names = F)
```

#### Synchrony 
Distribution of pairwise pearson correlations between tree ring series broken down by species and site (based on spline detrended series).
```{r synchrony, echo=F}
rings_field <- rings_field %>% 
  group_by(tree.uniqueID) %>% 
  mutate(minyear=min(Year)) %>% 
  ungroup() %>% as.data.frame()

for(i in unique(rings_field$Region)){
  tree_sub <- filter(rings_field, Region==i) %>% 
      filter(Year>=max(minyear)) %>%
      select(tree.uniqueID, Year, spline_growth) %>% 
      pivot_wider(names_from="tree.uniqueID", values_from="spline_growth") 
    
    if(dim(tree_sub)[1]==0){    
      cordf <- NA
    } else {
        treemat <-tree_sub %>% 
          select(-Year) %>%  as.matrix()

        cormat <- rcorr(treemat, type='pearson')$r
        cormat[upper.tri(cormat)] <- NA

       cordf <- melt(cormat) %>% filter(value!=1&!is.na(value)) %>% 
          rename(tree1=Var1, tree2=Var2, pearson_r=value) 
      }

    
    if(i=="San Jac"){all_cordf <- cordf
      } else all_cordf <- rbind(all_cordf, cordf)
}

sub <- select(rings_field, c(tree.uniqueID, Site, Species, Region, hilo, hegyi, DBH)) %>% 
  group_by(tree.uniqueID) %>% 
  summarise_all(first)

dist_mat2 <- select(dist_mat, c(focalID, compID, Distance)) %>% 
  rename(tree1=focalID, tree2=compID)

all_cordf2 <- all_cordf %>% 
  na.omit() %>% 
  mutate_at("tree1", as.character) %>% 
  left_join(sub, by=c("tree1"="tree.uniqueID")) %>% 
  left_join(select(sub, c(tree.uniqueID, Site, Species)), by=c("tree2"="tree.uniqueID")) %>% 
  left_join(dist_mat2)


write.csv(all_cordf2, "Processed Data/synchrony_dat.csv", row.names = F)

for(i in unique(rings_field$Region)){
  tree_sub <- filter(rings_field, Region==i) %>% 
      filter(Year>=max(minyear)) %>%
      select(tree.uniqueID, Year, raw_growth) %>% 
      pivot_wider(names_from="tree.uniqueID", values_from="raw_growth") 
    
    if(dim(tree_sub)[1]==0){    
      cordf <- NA
    } else {
        treemat <-tree_sub %>% 
          select(-Year) %>%  as.matrix()

        cormat <- rcorr(treemat, type='pearson')$r
        cormat[upper.tri(cormat)] <- NA

       cordf <- melt(cormat) %>% filter(value!=1&!is.na(value)) %>% 
          rename(tree1=Var1, tree2=Var2, pearson_r=value) 
      }

    
    if(i=="San Jac"){all_cordf <- cordf
      } else all_cordf <- rbind(all_cordf, cordf)
}


all_cordf_raw <- all_cordf %>% 
  na.omit() %>% 
  mutate_at("tree1", as.character) %>% 
  left_join(sub, by=c("tree1"="tree.uniqueID")) %>% 
  left_join(select(sub, c(tree.uniqueID, Site, Species)), by=c("tree2"="tree.uniqueID")) %>% 
  left_join(dist_mat2)

write.csv(all_cordf_raw, "Processed Data/synchrony_raw.csv", row.names = F)

```