---
title: "Synchrony from multivariate models"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
author: "Teresa Bohner"
date: "4/28/2020"
output: 
  html_document:
    toc: true
    collapsed: false
    toc_float: true
    toc_depth: 4
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
```

```{r packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(brms)
library(tidybayes)
library(viridis)
library(scales)
library(cowplot)
library(ggridges)
library(ggstance)
library(modelr)

```

## Overview

### Questions
1.  How does synchrony of tree growth vary across regional/elevation gradients?
Prediction: higher synchrony in lower latitude and lower elevation (drier) populations (SIASH, dendrochronological principles)

2.  Within populations is intraspecific synchrony greater than interspecific synchrony? Between specific species pairs? With increasing distance?
Prediction: higher synchrony among intraspecific pairs, potentially higher synchrony between pine pairs versus pine-fir pairs (successional stages), and lower synchrony with increasing distance

3.  How has synchrony changed through time? Are changes more dramatic in certain populations? Is increased or decreased synchrony associated with certain environmental variables?
Prediction: synchrony has increased with time, increased synchrony associated with drier, more variable time windows, changes are more pronounced in xeric populations.

Synchrony is explored here through residual correlations betweeen trees from multivariate models of tree growth.
```{r data}
precip <- read.csv("Processed Data/precip_temp_spei.csv")

timeseq <- seq(1910, 2010, by=10)

precip_summary <- expand_grid(precip, timeseq) %>% 
  filter(Year-timeseq>-11, Year-timeseq<11) %>% 
  group_by(Site, siteno, Neighborhood, hilo, timeseq) %>% 
  summarise_at(vars(total_ppt_mm, mean_temp_C, spei12), list(mean=mean, sd=sd))

rings_field <- read.csv("Processed Data/matched_rings_field.csv")  %>% 
  left_join(precip) %>% 
  group_by(Site, Neighborhood, Species) %>% 
  mutate(treeID=as.numeric(droplevels(as.factor(tree.uniqueID)))) %>% ## counter id if we want it rather than unique ID
  arrange(Site, Neighborhood, Species, ID) 
  

region <- rings_field %>% 
  ungroup() %>% 
  dplyr::select(c(Site, Region, hilo)) %>% 
  group_by(Site) %>% 
  summarize_all(first) %>% 
  mutate(Region=factor(Region, levels=c("San Jac", "SENF", "SEKI", "Mammoth")))

ids <- ungroup(rings_field) %>% 
  dplyr::select(c(Site, Neighborhood, tree.uniqueID)) %>% 
  group_by(Site, Neighborhood, tree.uniqueID) %>%  summarize_all(first)

bigvars2 <- read.csv("Processed Data/synchrony_multivar.csv")
# bigdraws2 <- read.csv("Processed Data/synchrony_multivar_draws.csv")
bigvars_t2 <- read.csv("Processed Data/synchrony_decadal.csv") 

sync_mod_dat <- read.csv("Processed Data/data_for_sync_mod.csv")

all_cor <- read.csv("Processed Data/inta_inter_pairwise_all.csv") %>% 
  mutate(pair2 =str_c(intra_spp,inter_spp, sep="-"))
  
```

## Q2 Intra-interspecific competition

We modeled the annual growth increment of all of the trees in a given neighborhood jointly to estimate their residual correlations (similar to calculating pearsons pairwise correlations).

### Pairwise residual correlation summaires: 
``` {r funs_and_manip}
## function to extract data at the species level----
spp_subsetter <- function(data, sp) {
  a <- filter(data, pair==paste(sort(c(sp, "ac")), collapse="-")) %>%
    mutate(focal_sp=sp,
           comp_sp="ac")
  b <- filter(data, pair==paste(sort(c(sp, "pj")), collapse="-")) %>%
    mutate(focal_sp=sp,
           comp_sp="pj")
  c <- filter(data, pair==paste(sort(c(sp, "pl")), collapse="-")) %>%
    mutate(focal_sp=sp,
           comp_sp="pl")
  d <- filter(data, pair==paste(sort(c(sp, "pp")), collapse="-")) %>%
    mutate(focal_sp=sp,
           comp_sp="pp")
  
  bind_rows(a,b,c,d) %>%
    mutate(comp_sp=ifelse(comp_sp==focal_sp, "self", comp_sp))
}

## summary data for median correlation
summary <- group_by(bigvars2, spp1, spp2, comp) %>% 
  summarise(med=median(.value))

## subset species----
acdat <- spp_subsetter(bigvars2, "ac")
pjdat <- spp_subsetter(bigvars2, "pj")
pldat <- spp_subsetter(bigvars2, "pl")
ppdat <- spp_subsetter(bigvars2, "pp")


# acdat <- spp_subsetter(bigdraws2, "ac")
# pjdat <- spp_subsetter(bigdraws2, "pj")
# pldat <- spp_subsetter(bigdraws2, "pl")
# ppdat <- spp_subsetter(bigdraws2, "pp")

```

*note difference in distribution with when keeping all of the draws versus summarizing median correlations. 


```{r fig}

## matrix style figure
ggplot(bigvars2, aes(.value, color=comp, fill=comp)) +
  geom_density(position = "identity", alpha=0.4) +
  theme_test() +
  geom_vline(data=summary, aes(xintercept = med), linetype="dashed") +
  facet_grid(spp2~spp1)

# ggplot(bigdraws2, aes(.value, color=comp, fill=comp)) +
#   geom_density(position = "identity", alpha=0.4) +
#   theme_test() +
#   geom_vline(data=summary, aes(xintercept = med), linetype="dashed") +
#   facet_grid(spp2~spp1)

# show_col(viridis_pal()(5)) # get the hex codes for the colors

cols <- c("self"="#440154FF", "pj"="#3B528BFF", "pl"="#21908CFF", "pp"="#5DC863FF", "ac"="#FDE725FF") ## keep species cols consistent

p1 <- ggplot(acdat, aes(.value, color=comp_sp, fill=comp_sp)) +
  geom_density(position = "identity", alpha=0.5) +
  theme_test() +
  ggtitle("A. concolor") +
  scale_fill_manual(values=cols) +
  scale_color_manual(values=cols) +
  theme(legend.position = 'none') +
  xlab("Correlation")

# acdat2 <- left_join(acdat, region) 
# 
# ggplot(acdat2, aes(.value, color=comp_sp, fill=comp_sp)) +
#   geom_density(position = "identity", alpha=0.5) +
#   theme_test() +
#   ggtitle("A. concolor") +
#   scale_fill_manual(values=cols) +
#   scale_color_manual(values=cols) +
#   theme(legend.position = 'none') +
#   facet_grid(Region~hilo)+
#   xlab("Correlation")

p2 <- ggplot(pjdat, aes(.value, color=comp_sp, fill=comp_sp)) +
  geom_density(position = "identity", alpha=0.5) +
  theme_test() +
  ggtitle("P. jeffreyi") +
  scale_fill_manual(values=cols) +
  scale_color_manual(values=cols) +
  theme(legend.position = 'none') +
  xlab("Correlation")

p3 <- ggplot(pldat, aes(.value, color=comp_sp, fill=comp_sp)) +
  geom_density(position = "identity", alpha=0.5) +
  theme_test() +
  ggtitle("P. lambertiana") +
  scale_fill_manual(values=cols) +
  scale_color_manual(values=cols) +
  theme(legend.position = 'none') +
  xlab("Correlation")

p4 <- ggplot(ppdat, aes(.value, color=comp_sp, fill=comp_sp)) +
  geom_density(position = "identity", alpha=0.5) +
  theme_test() +
  ggtitle("P. ponderosa") +
  scale_fill_manual(values=cols) +
  scale_color_manual(values=cols) +
  theme(legend.position = 'none') +
  xlab("Correlation")

dummy <- ggplot(bind_rows(acdat, pjdat), aes(.value, color=comp_sp, fill=comp_sp)) +
  geom_density(alpha=0.5) +
  theme_test() +
  ggtitle("P. lambertiana") +
  scale_fill_manual(values=cols) +
  scale_color_manual(values=cols) +
  labs(fill = "Competitor", color="Competitor")

legend <- cowplot::get_legend(dummy)
# plot_grid(legend)

plot_grid(plot_grid(p1,p2,p3,p4, ncol=2), legend, ncol=2, rel_widths = c(1,.2))
```

### Correlation response model: 
We wanted to test the relationship between the pairwise correlations and distance, size ratios, species-pairs, and locations of all of the trees. We built hierarchical mixed models with the pairwise correlations as a responses (normal). 

```{r sync_mixed_model}
pair_mod10 <- readRDS("saved models/sync_mods/pair_mod10.rds")
  
dat_post <- pair_mod10 %>% 
  spread_draws(b_Intercept, `r_Site:pair`[comp,]) %>% 
  mutate(comp_med=b_Intercept + `r_Site:pair`) %>% 
  dplyr::select(-c(b_Intercept, `r_Site:pair`)) %>% 
  median_qi() %>% 
  mutate(Site=str_sub(comp, 1,2),
         pair=str_sub(comp, 4,8),
         spp1=str_sub(pair, 1,2),
         spp2=str_sub(pair, 4,5)) %>% 
  left_join(region, by="Site")


ggplot(dat_post, aes(y=interaction(Region, hilo, lex.order=TRUE), x=comp_med, color=Region)) +
  geom_pointintervalh(position = position_dodgev(-0.4)) +
  scale_color_viridis(discrete=T, option='plasma', direction = -1) +
  facet_grid(spp1~spp2) +
  theme_test() +
  scale_x_continuous(labels = scales::number_format(accuracy = 0.1)) +
  xlab("posterior median intercept") +
  ylab("Site") +
  theme(legend.position = 'none')


pred_dat <- sync_mod_dat %>% 
  group_by(Site, pair) %>% 
  data_grid(cor, dist=round(quantile(sync_mod_dat$dist, probs=c(0.5)), digits=2), 
            sizeratio=round(quantile(sync_mod_dat$sizeratio, probs=c(0.5)), digits=2)) %>% 
  add_fitted_draws(pair_mod10, n=100, re_formula = NULL, allow_new_levels=TRUE) %>% 
  mutate(spp1=str_sub(pair, 1,2),
         spp2=str_sub(pair, 4,5),
         .value=.value) %>% 
  left_join(region, by="Site") 

## check out predicted data 
ggplot(pred_dat, aes(y = interaction(Region, hilo, lex.order=TRUE), x = .value, color=Region)) +
  stat_pointintervalh(.width = 0.95) +
  # geom_violinh(aes(x = .value)) +
  geom_vline(xintercept=0, linetype="dashed")+
  facet_grid(spp1~spp2) +
  scale_color_viridis(discrete=TRUE, option='plasma', direction = -1) +
  geom_hline(yintercept = 0, linetype="dashed") +
  # ylim(-.4, .4) +
  theme_test() +
  xlab("conditional effect")  +
  ylab("Site") +
  theme(legend.position = 'none')

dat_post <- pair_mod10 %>% 
  gather_draws(b_dist, b_sizeratio)

params <- ggplot(dat_post, aes(x=.value, y=.variable)) +
  geom_halfeyeh() +
  geom_vline(xintercept=0, linetype="dashed") +
  theme_test() +
  ylab("parameter") +
  xlab("estimate")

 pred_dat <- sync_mod_dat %>% 
  group_by(Site, pair) %>% 
  data_grid(cor, dist=round(quantile(sync_mod_dat$dist, probs=c(0.025, 0.25, 0.5, 0.75, 0.975)), digits=2), 
            sizeratio=round(quantile(sync_mod_dat$sizeratio, probs=c(0.5)), digits=2)) %>% 
  add_fitted_draws(pair_mod10, n=100, re_formula = NULL, allow_new_levels=TRUE) %>% 
  mutate(.value=.value)
  
dist <- ggplot(pred_dat, aes(x = dist, y = .value)) +
  stat_lineribbon(aes(y = .value), .width = c(.95), alpha=0.2) +
  stat_lineribbon(aes(y = .value), .width = c(.01)) +
  scale_fill_brewer(palette = "Greys") +
  scale_color_viridis(discrete=TRUE) +
  # ylim(-.4, .4) +
  theme_test() +
  ylab("predicted correlation (Rij)") +
  theme(legend.position = 'none') +
  xlab("Distance")

 pred_dat <- sync_mod_dat %>% 
  group_by(Site, pair) %>% 
  data_grid(cor, dist=round(quantile(sync_mod_dat$dist, probs=c(0.5)), digits=2), 
            sizeratio=round(quantile(sync_mod_dat$sizeratio, probs=c(0.025, 0.25, 0.5, 0.75, 0.975)), digits=2)) %>% 
  add_fitted_draws(pair_mod10, n=100, re_formula = NULL, allow_new_levels=TRUE)  %>% 
  mutate(.value=.value)
  
ratio <- ggplot(pred_dat, aes(x = sizeratio, y = .value)) +
  stat_lineribbon(aes(y = .value), .width = c(.95), alpha=0.2) +
  stat_lineribbon(aes(y = .value), .width = c(.01)) +
  scale_fill_brewer(palette = "Greys") +
  scale_color_viridis(discrete=TRUE) +
  # ylim(-.4, .4) +
  theme_test() +
  ylab("predicted correlation (Rij)")  +
  theme(legend.position = 'none') +
  xlab("Size Ratio")

plot_grid(params, plot_grid(dist, ratio, nrow=1), nrow=2)

```

### Fraction intraspecific > interspecific model: 
Next we wanted to know if the intraspecific correlations were significantly larger than the interspecific correlations and how much the species pair and/or geographic location mattered. Bernoulli error distribution with individual as group effect (as opposed to doing an aggregated binomial model...does this make sense?).  Data is in the form of a matrix for every individual tree of the comparison between the correlations between it and all of it's intraspecific neighbors and all of the interspecific neighbors.

```{r fraction_intra_model}
bern_mod9 <- readRDS("saved models/intra_greater_mods/bern_mod9.rds")
head(dplyr::select(all_cor, c(tree1.x, tree2.x, comp.x, .value.x, tree1.y, tree2.y, comp.y, .value.y, intra_greater)))

dat_post <- bern_mod9 %>% 
  spread_draws(b_Intercept, `r_Site:intra_spp:pair2`[comp,]) %>% 
  mutate(comp_med=inv_logit_scaled(b_Intercept + `r_Site:intra_spp:pair2`)) %>% 
  dplyr::select(-c(b_Intercept, `r_Site:intra_spp:pair2`)) %>% 
  median_qi() %>% 
  # mutate(Region=str_split_fixed(comp, "_", 3)[,1],
  #        intra_spp=str_sub(str_split_fixed(comp, "_", 3)[,3], 1,2),
  #        inter_spp=str_sub(str_split_fixed(comp, "_", 3)[,3], 4,5),
  #        Region=str_replace(Region, "\\.", " ")) %>% 
  mutate(Site=str_sub(comp, 1,2),
         intra_spp=str_sub(str_split_fixed(comp, "_", 3)[,3], 1,2),
         inter_spp=str_sub(str_split_fixed(comp, "_", 3)[,3], 4,5)) %>% 
  left_join(region, by="Site")


ggplot(dat_post, aes(y=reorder(intra_spp, desc(intra_spp)), x=comp_med, 
                     color=interaction(Region, hilo, lex.order=TRUE))) +
  geom_pointintervalh(position = position_dodgev(-0.4)) +
  scale_color_viridis(discrete=T, option='plasma', direction = -1) +
  geom_vline(xintercept=0.5, linetype="dashed") +
  facet_grid(~inter_spp) +
  theme_test() +
  scale_x_continuous(labels = scales::number_format(accuracy = 0.01)) +
  xlab("posterior median proportion of intra > inter correlations") +
  ylab("Intraspecific species pair") +
  theme(legend.position = 'none')

pred_dat <- all_cor %>% 
  group_by(Site, intra_spp, pair2) %>% 
  data_grid(intra_greater) %>% 
  add_fitted_draws(bern_mod9, re_formula = ~(1|Site/intra_spp/pair2), allow_new_levels=TRUE) %>% 
  mutate(.value=inv_logit_scaled(.value),
         spp1=str_sub(pair2, 1,2),
         spp2=str_sub(pair2, 4,5))

## check out predicted data 
ggplot(pred_dat, aes(x = spp2, y = .value, fill=spp2)) +
  geom_violin(aes(y = .value)) +
  facet_grid(~spp1) +
  scale_fill_viridis(discrete=TRUE) +
  # scale_alpha_ordinal(range=c(0.3,1)) +
  geom_hline(yintercept = 0.5, linetype="dashed") +
  # ylim(-.4, .4) +
  theme_test() +
  ylab("predicted proportion of intra > inter correlations") 

```

## Q3 Synchrony through time
Used 20 year time blocks overlapping 10 years



```{r time_manip}
## subset species----
acdat <- spp_subsetter(bigvars_t2, "ac")
pjdat <- spp_subsetter(bigvars_t2, "pj")
pldat <- spp_subsetter(bigvars_t2, "pl")
ppdat <- spp_subsetter(bigvars_t2, "pp")

cols <- c("self"="#440154FF", "pj"="#3B528BFF", "pl"="#21908CFF", "pp"="#5DC863FF", "ac"="#FDE725FF") ## keep species cols consistent

ggplot(acdat, aes(x=.value, y=as.factor(time), fill=comp_sp)) +
  stat_density_ridges(quantile_lines = TRUE, quantiles = 2,  linetype="dashed") +
  stat_density_ridges(alpha=0) +
  facet_grid(~comp_sp) +
  theme_test() +
  ggtitle("A. concolor") +
  scale_fill_manual(values=cols) +
  theme(legend.position = 'none') +
  xlab("Correlation") +
  ylab("20 year block (year +/- 10)")

ggplot(pjdat, aes(x=.value, y=as.factor(time), fill=comp_sp)) +
  stat_density_ridges(quantile_lines = TRUE, quantiles = 2,  linetype="dashed") +
  stat_density_ridges(alpha=0) +
  facet_grid(~comp_sp) +
  theme_test() +
  ggtitle("P. jeffreyi") +
  scale_fill_manual(values=cols) +
  theme(legend.position = 'none') +
  xlab("Correlation") +
  ylab("20 year block (year +/- 10)")

ggplot(pldat, aes(x=.value, y=as.factor(time), fill=comp_sp)) +
  stat_density_ridges(quantile_lines = TRUE, quantiles = 2,  linetype="dashed") +
  stat_density_ridges(alpha=0) +
  facet_grid(~comp_sp) +
  theme_test() +
  ggtitle("P. lambertiana") +
  scale_fill_manual(values=cols) +
  theme(legend.position = 'none') +
  xlab("Correlation") +
  ylab("20 year block (year +/- 10)")

ggplot(ppdat, aes(x=.value, y=as.factor(time), fill=comp_sp)) +
  stat_density_ridges(quantile_lines = TRUE, quantiles = 2,  linetype="dashed") +
  stat_density_ridges(alpha=0) +
  facet_grid(~comp_sp) +
  theme_test() +
  ggtitle("P. ponderosa") +
  scale_fill_manual(values=cols) +
  theme(legend.position = 'none') +
  xlab("Correlation") +
  ylab("20 year block (year +/- 10)")

```

```{r climate_time}
# ggplot(precip_summary, aes(timeseq, total_ppt_mm_sd, color=Site)) +
#   geom_point() 
# 
# ggplot(precip_summary, aes(timeseq, total_ppt_mm_mean, color=Site)) +
#   geom_point()

ggplot(precip_summary, aes(total_ppt_mm_mean, total_ppt_mm_sd, color=timeseq)) +
  geom_point() +
  theme_test() 

# 
# ggplot(precip_summary, aes(timeseq, mean_temp_C_sd, color=Site)) +
#   geom_point()
# 
# ggplot(precip_summary, aes(timeseq, mean_temp_C_mean, color=Site)) +
#   geom_point()

ggplot(precip_summary, aes(mean_temp_C_mean, mean_temp_C_sd, color=timeseq)) +
  geom_point()+
  theme_test() 

# ggplot(precip_summary, aes(timeseq, spei12_sd, color=Site)) +
#   geom_point()
# 
# ggplot(precip_summary, aes(timeseq, spei12_mean, color=Site)) +
#   geom_point()

ggplot(precip_summary, aes(spei12_mean, spei12_sd, color=timeseq)) +
  geom_point()+
  theme_test() 


```



```{r time_models}
alldata <- bigvars_t2 %>% 
  select(.value, tree1, tree2, spp1, spp2, pair, comp, Site, Neighborhood, time) %>% 
  rename(corr=.value) %>% 
  left_join(precip_summary, by=c("Site","Neighborhood", "time"="timeseq"))

# mod <- brm(corr~time + (time|pair), data=alldata, cores=4)

```