---
title: "Synchrony analysis"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
author: "Teresa Bohner"
date: "4/28/2020"
output: 
  html_document:
    toc: true
    collapsed: false
    toc_float: true
    toc_depth: 4
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
knitr::opts_chunk$set(fig.path = "figures/")
```

```{r packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(brms)
library(tidybayes)
library(viridis)
library(scales)
library(cowplot)
library(ggridges)
library(ggstance)
library(modelr)

```

## Overview

### Questions
1.  How does synchrony of tree growth vary across regional/elevation gradients?
Prediction: higher synchrony in lower latitude and lower elevation (drier) populations (SIASH, dendrochronological principles)

2.  Within populations is intraspecific synchrony greater than interspecific synchrony? Between specific species pairs? With increasing distance?
Prediction: higher synchrony among intraspecific pairs, potentially higher synchrony between pine pairs versus pine-fir pairs (successional stages), and lower synchrony with increasing distance

3.  How has synchrony changed through time? Are changes more dramatic in certain populations? Is increased or decreased synchrony associated with certain environmental variables?
Prediction: synchrony has increased with time, increased synchrony associated with drier, more variable time windows, changes are more pronounced in xeric populations.

Synchrony is explored here through residual correlations betweeen trees from multivariate models of tree growth.
```{r data}
rings_field <- read.csv("Processed Data/matched_rings_field.csv")  %>% 
  group_by(Site, Neighborhood, Species) %>% 
  mutate(comp_BA=BA_con+BA_het) %>%
  arrange(Site, Neighborhood, Species, ID)
  
region <- rings_field %>% 
  ungroup() %>% 
  dplyr::select(c(Site, Region, hilo)) %>% 
  group_by(Site) %>% 
  summarize_all(first) %>% 
  mutate(Region=factor(Region, levels=c("San Jac", "SENF", "SEKI", "Mammoth")))

ids <- ungroup(rings_field) %>% 
  dplyr::select(c(Site, Neighborhood, tree.uniqueID)) %>% 
  group_by(Site, Neighborhood, tree.uniqueID) %>%  summarize_all(first)

tree_data <- rings_field %>% 
  group_by(tree.uniqueID) %>% 
  summarise(DBH=first(DBH),
            comp_BA=first(comp_BA))

precip <- read.csv("Processed Data/precip_temp_spei.csv")

decade <- seq(1910, 2020, by=10)

precip_summary <- expand_grid(precip, decade) %>% 
  filter(Year-decade>-29, Year<=decade) %>% 
  group_by(Site, siteno, Neighborhood, hilo, decade) %>% 
  summarise_at(vars(total_ppt_mm, mean_temp_C, spei12), list(mean=function(x) mean(x, na.rm=T), sd=function(x) sd(x, na.rm=T))) %>% 
  mutate_at(vars(total_ppt_mm_mean, mean_temp_C_mean,
                 total_ppt_mm_sd, mean_temp_C_sd), .funs = function(x) scale(x, scale = FALSE)) %>% 
  left_join(region) %>% 
  ungroup()

spp_pairs <- expand_grid(Species.x=c("ac", "pj", "pl", "pp"), Species.y=c("ac", "pj", "pl", "pp")) %>% 
  mutate(pair_first=str_c(Species.x, Species.y, sep="-"),
         pair=c("ac-ac", "ac-pj", "ac-pl", "ac-pj", "ac-pj", "pj-pj", "pj-pl", "pj-pp", 
                      "ac-pl", "pj-pl", "pl-pl", "pl-pp", "ac-pp", "pj-pp", "pl-pp", "pp-pp")) %>% 
  dplyr::select(-c(pair_first))

pearson <- read.csv("Processed Data/synchrony_dat.csv") %>% 
  na.omit() %>% 
  filter(Site.x !="ic") %>% 
  mutate_at(vars(Species.x, Species.y), tolower) %>% 
  left_join(spp_pairs) %>% 
  mutate(spp1=str_sub(pair, 1,2),
         spp2=str_sub(pair, 4,5),
         comp=ifelse(Species.x==Species.y, "intra", "inter"))


pearson_t <- read.csv("Processed Data/synchrony_decadal_pearson.csv") %>% 
  na.omit() %>% 
  filter(Site.x !="ic") %>% 
  mutate_at(vars(Species.x, Species.y), tolower) %>% 
  left_join(spp_pairs) %>% 
  mutate(spp1=str_sub(pair, 1,2),
         spp2=str_sub(pair, 4,5),
         comp=ifelse(Species.x==Species.y, "intra", "inter"))
summary_t <- pearson_t %>% 
  dplyr::select(c(pearson_r, decade, Site.x, Neighborhood.x, pair, comp)) %>% 
  group_by(Site.x, Neighborhood.x, pair, comp, decade) %>% #
  summarise(samp.depth=length(pearson_r),
            pearson_sd=sd(pearson_r, na.rm=T),
            pearson_r=mean(pearson_r, na.rm=T)) %>% 
  left_join(precip_summary, by=c("Site.x"="Site", "Neighborhood.x"="Neighborhood", "decade")) #

## join all data ----
alldata <- pearson %>% 
  left_join(tree_data, by=c("tree2"="tree.uniqueID")) %>% 
  mutate(sizeratio=ifelse(DBH.x>DBH.y, DBH.x/DBH.y, DBH.y/DBH.x),
         Region=factor(Region, levels=c("San Jac", "SENF", "SEKI", "Mammoth"))) %>% 
  mutate_at(vars(dist, sizeratio, comp_BA), scale)

alldata_t <- pearson_t %>% 
  left_join(tree_data, by=c("tree2"="tree.uniqueID")) %>% 
  mutate(sizeratio=ifelse(DBH.x>DBH.y, DBH.x/DBH.y, DBH.y/DBH.x),
         Region=factor(Region, levels=c("San Jac", "SENF", "SEKI", "Mammoth"))) %>% 
  mutate_at(vars(dist, sizeratio, comp_BA), scale) %>% 
  left_join(dplyr::select(precip_summary, -c(Region, hilo)), by=c('Site.x'='Site', 'decade'='decade'))

  
```

## Q2 Intra-interspecific competition

### Pairwise pearson correlation summaires: 
We calculated the correlation between series of tree ring growth for all unique pairs of individuals within competitive neighborhoods. 
``` {r funs_and_manip}
## function to extract data at the species level----
spp_subsetter <- function(data, sp) {
  a <- filter(data, pair==paste(sort(c(sp, "ac")), collapse="-")) %>%
    mutate(focal_sp=sp,
           comp_sp="ac")
  b <- filter(data, pair==paste(sort(c(sp, "pj")), collapse="-")) %>%
    mutate(focal_sp=sp,
           comp_sp="pj")
  c <- filter(data, pair==paste(sort(c(sp, "pl")), collapse="-")) %>%
    mutate(focal_sp=sp,
           comp_sp="pl")
  d <- filter(data, pair==paste(sort(c(sp, "pp")), collapse="-")) %>%
    mutate(focal_sp=sp,
           comp_sp="pp")
  
  bind_rows(a,b,c,d) %>%
    mutate(comp_sp=ifelse(comp_sp==focal_sp, "self", comp_sp))
}

## summary data for median correlation
summary <- group_by(alldata, spp1, spp2, pair) %>% 
  summarise(med=median(pearson_r))

## subset species----
acdat <- spp_subsetter(alldata, "ac")
pjdat <- spp_subsetter(alldata, "pj")
pldat <- spp_subsetter(alldata, "pl")
ppdat <- spp_subsetter(alldata, "pp")


```



```{r fig}
## matrix style figure
ggplot(alldata, aes(pearson_r, fill=comp)) +
  geom_density(aes(y = ..scaled..), alpha=0.5) +
  facet_grid(spp2~spp1) +
  theme_test() +
  geom_vline(data=summary, aes(xintercept=med), linetype="dashed") +
  ylab("proportion") + xlab("pearson correlation")

# show_col(viridis_pal()(5)) # get the hex codes for the colors
#440154FF
cols <- c("self"="#440154FF", "pj"="#3B528BFF", "pl"="#21908CFF", "pp"="#5DC863FF", "ac"="#FDE725FF") ## keep species cols consistent

p1 <- ggplot(acdat, aes(pearson_r, color=comp_sp, fill=comp_sp)) +
  geom_density(position = "identity", alpha=0.5) +
  theme_test() +
  ggtitle("A. concolor") +
  scale_fill_manual(values=cols) +
  scale_color_manual(values=cols) +
  theme(legend.position = 'none') +
  xlab("pearson correlation")

p2 <- ggplot(pjdat, aes(pearson_r, color=comp_sp, fill=comp_sp)) +
  geom_density(position = "identity", alpha=0.5) +
  theme_test() +
  ggtitle("P. jeffreyi") +
  scale_fill_manual(values=cols) +
  scale_color_manual(values=cols) +
  theme(legend.position = 'none') +
  xlab("pearson correlation")

p3 <- ggplot(pldat, aes(pearson_r, color=comp_sp, fill=comp_sp)) +
  geom_density(position = "identity", alpha=0.5) +
  theme_test() +
  ggtitle("P. lambertiana") +
  scale_fill_manual(values=cols) +
  scale_color_manual(values=cols) +
  theme(legend.position = 'none') +
  xlab("pearson correlation")

p4 <- ggplot(ppdat, aes(pearson_r, color=comp_sp, fill=comp_sp)) +
  geom_density(position = "identity", alpha=0.5) +
  theme_test() +
  ggtitle("P. ponderosa") +
  scale_fill_manual(values=cols) +
  scale_color_manual(values=cols) +
  theme(legend.position = 'none') +
  xlab("pearson correlation")

dummy <- ggplot(bind_rows(acdat, pjdat), aes(pearson_r, color=comp_sp, fill=comp_sp)) +
  geom_density(alpha=0.5) +
  theme_test() +
  ggtitle("P. lambertiana") +
  scale_fill_manual(values=cols) +
  scale_color_manual(values=cols) +
  labs(fill = "Competitor", color="Competitor")

legend <- cowplot::get_legend(dummy)
# plot_grid(legend)

plot_grid(plot_grid(p1,p2,p3,p4, ncol=2), legend, ncol=2, rel_widths = c(1,.2))
```

### Correlation response model: 
We wanted to test the relationship between the pairwise correlations and distance, size ratios, species-pairs, and locations of all of the trees. We built hierarchical mixed models with the pairwise correlations as a responses (normal truncated response (-1,1)). 

```{r sync_mixed_model}
mod12 <- readRDS("saved models/pearson models/mod12.rds")
print(mod12$formula)

mod12 %>% 
  gather_draws(b_dist, b_sizeratio) %>% 
  median_qi(.width = c(.50, .95)) %>% 
  ggplot(aes(y=.variable, x=.value)) +
  geom_pointintervalh() +
  geom_vline(xintercept=0, linetype="dashed") +
  theme_test() + ylab("") + xlab("parameter estimate") +
  labs(caption="Fig. Increasing distance and larger differences between tree sizes correspond with lower synchrony values") +
  theme(plot.caption = element_text(hjust = 0))

mod12 %>% 
  spread_draws(b_Intercept, r_pair[pair,]) %>% 
  mutate(spp_mean=b_Intercept + r_pair) %>% 
  median_qi(.width = c(.50, .95)) %>% 
  ggplot(aes(y=pair, x=spp_mean, xmin=spp_mean.lower, xmax=spp_mean.upper)) +
  geom_pointintervalh() +
  geom_vline(xintercept=0, linetype="dashed") +
  theme_test() + ylab("") + xlab("Synchrony effect estimate") +
  labs(caption="Fig. Variation in effect of species pairing on synchrony") +
  theme(plot.caption = element_text(hjust = 0))


mod12 %>% 
  spread_draws(r_pair[pair,]) %>% 
  compare_levels(r_pair, by = pair) %>%
  median_qi(.width = c(.50, .95)) %>% 
  mutate(pair1=str_sub(pair, 1,5),
         pair1_spp1=str_sub(pair, 1,2),
         pair1_spp2=str_sub(pair, 4,5),
         pair1_comp=ifelse(pair1_spp1==pair1_spp2, "intra", "inter"),
         pair2=str_sub(pair, 9,13),
         pair2_spp1=str_sub(pair, 9,10),
         pair2_spp2=str_sub(pair, 12,13),
         pair2_comp=ifelse(pair2_spp1==pair2_spp2, "intra", "inter"),
         valid=ifelse(pair1_comp=="intra", str_detect(pair2, pair1_spp1), str_detect(pair1, pair2_spp1))) %>% 
  filter(pair1_comp!=pair2_comp, valid==TRUE) %>% 
  mutate(new_pair=ifelse(pair1_comp=="intra", str_c(pair1, pair2, sep=" - "), str_c(pair2, pair1, sep=" - ")),
         r_pair=ifelse(pair1_comp=="intra", r_pair, r_pair*-1),
         .upper=ifelse(pair1_comp=="intra", .upper, .upper*-1),
         .lower=ifelse(pair1_comp=="intra", .lower, .lower*-1)) %>% 
  ggplot(aes(y = new_pair, x = r_pair)) +
  geom_pointintervalh() +
  geom_vline(xintercept=0, linetype="dashed") +
  theme_test() + ylab("Species pair comparisons") + xlab("Synchrony contrast estimate") +
  labs(caption="Fig. Contrast in intra versus interspecific synchrony effect. Positive contrasts indicate that intraspecific \n synchrony is greater than interspecific. White fir and sugar pine show synchrony differentiation but not  \nthe yellow pines") +
  theme(plot.caption = element_text(hjust = 0))

mod12 %>% 
  spread_draws(b_Intercept, r_pair[pair,], `r_pair:Region:hilo`[site,]) %>% 
  mutate(spp_mean=b_Intercept + r_pair + `r_pair:Region:hilo`,
         valid=str_detect(site, pair)) %>% 
  filter(valid==1) %>% 
  median_qi(.width = c(.50, .95)) %>%  
  mutate(newsite=str_sub(site, start=7)) %>% 
  ggplot(aes(y=pair, x=spp_mean, xmin=spp_mean.lower, xmax=spp_mean.upper, color=newsite)) +
  geom_pointintervalh(position = position_dodgev(height=0.5)) +
  geom_vline(xintercept=0, linetype="dashed") +
  theme_test() + ylab("") + xlab("Synchrony effect estimate") +
  labs(caption="Fig. Variation in effect of species pairing on synchrony including region and elevation variation") +
  theme(plot.caption = element_text(hjust = 0))

mod12 %>% 
  spread_draws(r_pair[pair,], `r_pair:Region:hilo`[site,]) %>% 
  mutate(spp_mean=r_pair + `r_pair:Region:hilo`,
         valid=str_detect(site, pair)) %>% 
  filter(valid==1) %>% 
  mutate(newsite=str_sub(site, start=7)) %>% 
  group_by(newsite) %>% 
  compare_levels(spp_mean, by = pair) %>%
  median_qi(.width = c(.50, .95)) %>%  
  mutate(pair1=str_sub(pair, 1,5),
         pair1_spp1=str_sub(pair, 1,2),
         pair1_spp2=str_sub(pair, 4,5),
         pair1_comp=ifelse(pair1_spp1==pair1_spp2, "intra", "inter"),
         pair2=str_sub(pair, 9,13),
         pair2_spp1=str_sub(pair, 9,10),
         pair2_spp2=str_sub(pair, 12,13),
         pair2_comp=ifelse(pair2_spp1==pair2_spp2, "intra", "inter"),
         valid=ifelse(pair1_comp=="intra", str_detect(pair2, pair1_spp1), str_detect(pair1, pair2_spp1))) %>% 
  filter(pair1_comp!=pair2_comp, valid==TRUE) %>% 
  mutate(new_pair=ifelse(pair1_comp=="intra", str_c(pair1, pair2, sep=" - "), str_c(pair2, pair1, sep=" - ")),
         spp_mean=ifelse(pair1_comp=="intra", spp_mean, spp_mean*-1),
         .upper=ifelse(pair1_comp=="intra", .upper, .upper*-1),
         .lower=ifelse(pair1_comp=="intra", .lower, .lower*-1)) %>% 
  ggplot(aes(y = new_pair, x = spp_mean, color=newsite)) +
  geom_pointintervalh(position = position_dodgev(height=0.5)) +
  geom_vline(xintercept=0, linetype="dashed") +
  theme_test() + ylab("Species pair comparisons") + xlab("Synchrony contrast estimate") +
  labs(caption="Fig. Contrast in intra versus interspecific synchrony effect including region and elevation variation. Positive \n contrasts indicate thatintraspecific synchrony is greater than interspecific. Species level patterns in differentiation \n predominate although there is noteworthy variation among sites.") +
  theme(plot.caption = element_text(hjust = 0))

```


## Q3 Synchrony through time
### Pairwise pearson correlations through time:
We calculated pearson correlations between pairs of individuals within neighborhoods for 30 year time blocks with a 10 year lag to measure synchrony through time. Trends were analyzed for data summarized for species-pairs at the neighborhood level. Note this is measuring within population synchrony (in contrast to Shestakova spatial synchrony).


```{r exploratory}
ggplot(alldata_t, aes(decade, pearson_r, color=Region)) +
  stat_summary(aes(group=interaction(Region, hilo)), geom="pointrange") +
  stat_summary(fun.y=mean, aes(group=interaction(Region, hilo), linetype=hilo), geom="line") +
  theme_test() +
  labs(caption="Fig. Growth synchrony through time.") +
  theme(plot.caption = element_text(hjust = 0))

ggplot(alldata_t, aes(decade, pearson_r, color=pair)) +
  stat_summary(aes(group=interaction(pair, hilo)), geom="pointrange") +
  stat_summary(fun.y=mean, aes(group=interaction(pair, hilo), linetype=hilo), geom="line") +
  theme_test() +
  facet_grid(Region~hilo) +
  labs(caption="Fig. Patterns in growth synchrony through time appear to be driven more by site than species pairing.") +
  theme(plot.caption = element_text(hjust = 0))

a <- ggplot(precip_summary, aes(decade, spei12_mean, color=Region)) +
  geom_point() +
  stat_summary(fun.y=mean, aes(group=interaction(Region, hilo), linetype=hilo), geom="line") +
  theme_test() + theme(legend.position = 'none')

b <- ggplot(precip_summary, aes(decade, spei12_sd, color=Region)) +
  geom_point()  +
  stat_summary(fun.y=mean, aes(group=interaction(Region, hilo), linetype=hilo), geom="line") +
  theme_test() + theme(legend.position = 'none')

# now add the title
caption <- ggdraw() + 
  draw_label(
    "Fig. mean annual SPEI and SD for the 30 year blocks. ",
    x = 0,
    hjust = 0
  ) +
  theme(
    # add margin on the left of the drawing canvas,
    # so title is aligned with left edge of first plot
    plot.margin = margin(0, 0, 0, 7)
  )
plot_grid(
  plot_grid(a, b, ncol=2, labels = "auto"), caption, 
  ncol = 1,
  # rel_heights values control vertical title margins
  rel_heights = c(1, .1)
)


a <- ggplot(precip_summary, aes(decade, total_ppt_mm_sd, color=Region)) +
  geom_point()  +
  stat_summary(fun.y=mean, aes(group=interaction(Region, hilo), linetype=hilo), geom="line") +
  theme_test() + theme(legend.position = 'none')
b <- ggplot(precip_summary, aes(decade, total_ppt_mm_mean, color=Region)) +
  geom_point() +
  stat_summary(fun.y=mean, aes(group=interaction(Region, hilo), linetype=hilo), geom="line") +
  theme_test() + theme(legend.position = 'none')

caption <- ggdraw() + 
  draw_label(
    "Fig. mean total annual precipitation (mm) and SD for the 30 year blocks. ",
    x = 0,
    hjust = 0
  ) +
  theme(
    # add margin on the left of the drawing canvas,
    # so title is aligned with left edge of first plot
    plot.margin = margin(0, 0, 0, 7)
  )
plot_grid(
  plot_grid(a, b, ncol=2, labels = "auto"), caption, 
  ncol = 1,
  # rel_heights values control vertical title margins
  rel_heights = c(1, .1)
)


a <- ggplot(precip_summary, aes(decade, mean_temp_C_mean, color=Region)) +
  geom_point() +
  stat_summary(fun.y=mean, aes(group=interaction(Region, hilo), linetype=hilo), geom="line") +
  theme_test() + theme(legend.position = 'none')

b <- ggplot(precip_summary, aes(decade, mean_temp_C_sd, color=Region)) +
  geom_point()  +
  stat_summary(fun.y=mean, aes(group=interaction(Region, hilo), linetype=hilo), geom="line") +
  theme_test() + theme(legend.position = 'none')

caption <- ggdraw() + 
  draw_label(
    "Fig. mean annual temp (C) and SD for the 30 year blocks. ",
    x = 0,
    hjust = 0
  ) +
  theme(
    # add margin on the left of the drawing canvas,
    # so title is aligned with left edge of first plot
    plot.margin = margin(0, 0, 0, 7)
  )
plot_grid(
  plot_grid(a, b, ncol=2, labels = "auto"), caption, 
  ncol = 1,
  # rel_heights values control vertical title margins
  rel_heights = c(1, .1)
)


# ggplot(summary_t, aes(spei12_mean, pearson_r, color=comp)) +
#   geom_point() + facet_grid(Region~hilo) +
#   geom_smooth(method='lm')
# 
# ggplot(summary_t, aes(spei12_sd, pearson_r, color=comp)) +
#   geom_point() + facet_grid(Region~hilo) +
#   geom_smooth(method='lm')
# 
# ggplot(summary_t, aes(mean_temp_C_mean, pearson_r, color=pair)) +
#   geom_point() + facet_grid(Region~hilo) +
#   geom_smooth(method='lm')
# 
# ggplot(summary_t, aes(total_ppt_mm_mean, pearson_r, color=comp)) +
#   geom_point() + facet_grid(Region~hilo) +
#   geom_smooth(method='lm')

```


### Spline model:
I wanted to explore fitting a spline model to look at nonlinear changes in synchrony over time for different species-pairs and at different sites.
```{r spline_model}
spline_mod <- readRDS("saved models/spline_site.rds")
print(spline_mod$formula)
me <- marginal_effects(spline_mod)

ggplot(me$decade, aes(decade, estimate__)) +
  geom_ribbon(aes(ymin=lower__, ymax=upper__), alpha=0.3) +
   geom_line() + theme_test() + ylab("expected correlation") + 
  labs(caption="Fig. Overall synchrony has increased through time.") +
  theme(plot.caption = element_text(hjust = 0))

ggplot(me$`decade:Site.x`, aes(decade, estimate__)) +
  geom_ribbon(aes(ymin=lower__, ymax=upper__, fill=Site.x), alpha=0.3) +
  geom_line(aes(color=Site.x)) + theme_test() + ylab("expected correlation") + 
  labs(caption="Fig. Fluctuations are variable at the site level, there is a particularly steep increase for black mountain (San Jacinto).") +
  theme(plot.caption = element_text(hjust = 0))

spline_mod <- readRDS("saved models/spline.rds")
## should go with changes through time model
time_summ <- expand_grid(Site.x=unique(droplevels(summary_t$Site.x)), pair=c('AC-AC', 'PJ-AC', 'PJ-PJ', 'PL-AC', 'PL-PJ', 'PL-PL', 'PP-AC', 'PP-PJ', 'PP-PL', 'PP-PP'), decade=unique(summary_t$decade)) %>%
  add_fitted_draws(spline_mod, re_formula = NULL, allow_new_levels=TRUE) %>%
  compare_levels(.value, by=pair) %>% 
  median_qi(.width=c(0.5, 0.95))

time_summ2 <- time_summ %>% 
  mutate(pair1=str_sub(pair, 1,5),
         pair1_spp1=str_sub(pair, 1,2),
         pair1_spp2=str_sub(pair, 4,5),
         pair1_comp=ifelse(pair1_spp1==pair1_spp2, "intra", "inter"),
         pair2=str_sub(pair, 9,13),
         pair2_spp1=str_sub(pair, 9,10),
         pair2_spp2=str_sub(pair, 12,13),
         pair2_comp=ifelse(pair2_spp1==pair2_spp2, "intra", "inter"),
         valid=ifelse(pair1_comp=="intra", str_detect(pair2, pair1_spp1), str_detect(pair1, pair2_spp1))) %>% 
  filter(pair1_comp!=pair2_comp, valid==TRUE) %>% 
  mutate(new_pair=ifelse(pair1_comp=="intra", str_c(pair1, pair2, sep=" - "), str_c(pair2, pair1, sep=" - ")),
         .value=ifelse(pair1_comp=="intra", .value, .value*-1),
         .upper=ifelse(pair1_comp=="intra", .upper, .upper*-1),
         .lower=ifelse(pair1_comp=="intra", .lower, .lower*-1),
         timechunk=ifelse(decade<1970, "early", "late")) %>% 
  dplyr::select(c(Site.x, new_pair, .value, .upper, .lower, .width, timechunk)) %>% 
  group_by(Site.x, new_pair, timechunk, .width) %>% 
  summarise_all(mean)

ggplot(time_summ2, aes(x=.value, y=new_pair, color=timechunk)) +
  geom_pointintervalh(position = position_dodgev(height=0.5)) + theme_test() +
  geom_vline(xintercept = 0, linetype="dashed")  +
  xlab("expected correlation")+ ylab("Species contrast") + 
  labs(caption="Fig. Changes in synchrony differences between intra and interspecific pairing. Post 1970 \n white fir intraspecific synchrony has become stronger than interspecific synchrony.") +
  theme(plot.caption = element_text(hjust = 0))



```

### Climate response model:
Model selection supported SPEI as the best predictor for synchrony.  I included both mean and sd in one model although the variables are correlatedd (~.3).  I did not do multiple regression because of multicollinearity. 

```{r env_model}
clim_mod <- readRDS("saved models/pearson models/clim8c.rds")
print(clim_mod$formula)
  
clim_summ <- expand_grid(Region=unique(summary_t$Region), hilo=unique(summary_t$hilo), spei12_mean=quantile(summary_t$spei12_mean, probs=c(.025, .25, .5, .75, .975)), spei12_sd=quantile(summary_t$spei12_sd, probs=c(.5))) %>% 
  mutate(Site=paste0(Region,hilo)) %>% 
  filter(Site!="Mammothlow") %>% 
  add_fitted_draws(clim_mod, re_formula = ~(spei12_mean + spei12_sd | Region/hilo) ) %>% 
  median_qi()

ggplot(clim_summ, aes(spei12_mean, .value)) +
  geom_ribbon(aes(ymin=.lower, ymax=.upper, fill=Region, group=interaction(Region, hilo)), alpha=0.2) +
  geom_line(aes(color=Region , group=interaction(Region, hilo))) + theme_test() +
  ylab("expected correlation") + 
  labs(caption="Fig. Regional relationship between SPEI mean and expected synchrony. In the lowest latitudes synchrony is highest in \n dry periods while in the most mesic site synchrony is higher in the wet periods.") +
  theme(plot.caption = element_text(hjust = 0))

clim_summ <- expand_grid(Region=unique(summary_t$Region), hilo=unique(summary_t$hilo), spei12_mean=quantile(summary_t$spei12_mean, probs=c( .5)), spei12_sd=quantile(summary_t$spei12_sd, probs=c(.025, .25, .5, .75, .975))) %>% 
  mutate(Site=paste0(Region,hilo)) %>% 
  filter(Site!="Mammothlow") %>% 
  add_fitted_draws(clim_mod, re_formula = ~(spei12_mean + spei12_sd | Region/hilo) ) %>% 
  median_qi()

ggplot(clim_summ, aes(spei12_sd, .value)) +
  geom_ribbon(aes(ymin=.lower, ymax=.upper, fill=Region, group=interaction(Region, hilo)), alpha=0.2) +
  geom_line(aes(color=Region , group=interaction(Region, hilo))) + theme_test() +
  ylab("expected correlation") + 
  labs(caption="Fig. Regional relationship between SPEI SD and expected synchrony. In the lowest latitudes synchrony is highest in \n variable periods while in the most mesic site synchrony is higher in the more variable periods") +
  theme(plot.caption = element_text(hjust = 0))

clim_summ <- expand_grid(pair=unique(summary_t$pair), spei12_mean=quantile(summary_t$spei12_mean, probs=c(.025, .25, .5, .75, .975)), spei12_sd=quantile(summary_t$spei12_sd, probs=c(.5))) %>% 
  add_fitted_draws(clim_mod, re_formula = ~(1 | pair) ) %>% 
  median_qi()

ggplot(clim_summ, aes(spei12_mean, .value)) +
  geom_ribbon(aes(ymin=.lower, ymax=.upper, fill=pair), alpha=0.2) +
  geom_line(aes(color=pair)) + theme_test() +
  ylab("expected correlation") + 
  labs(caption="Fig. There appears to be little effect of species pair on the relationship between SPEI period mean and synchrony. ") +
  theme(plot.caption = element_text(hjust = 0))

clim_summ <- expand_grid(pair=unique(summary_t$pair), spei12_mean=quantile(summary_t$spei12_mean, probs=c( .5)), spei12_sd=quantile(summary_t$spei12_sd, probs=c(.025, .25, .5, .75, .975))) %>% 
  add_fitted_draws(clim_mod, re_formula = ~(1 | pair) ) %>% 
  median_qi()

ggplot(clim_summ, aes(spei12_sd, .value)) +
  geom_ribbon(aes(ymin=.lower, ymax=.upper, fill=pair), alpha=0.2) +
  geom_line(aes(color=pair)) + theme_test()  +
  ylab("expected correlation") + 
  labs(caption="Fig. Overall synchrony increases among all species pairs in more variable periods.") +
  theme(plot.caption = element_text(hjust = 0))

clim_summ <- expand_grid(Region=unique(summary_t$Region), hilo=unique(summary_t$hilo), pair=unique(summary_t$pair), spei12_mean=quantile(summary_t$spei12_mean, probs=c(.025, .25, .5, .75, .975)), spei12_sd=quantile(summary_t$spei12_sd, probs=c(.5))) %>% 
  mutate(Site=paste0(Region,hilo)) %>% 
  filter(Site!="Mammothlow") %>% 
  add_fitted_draws(clim_mod, re_formula = NULL, allow_new_levels=TRUE) %>% 
  median_qi()

ggplot(clim_summ, aes(spei12_mean, .value)) +
  geom_ribbon(aes(ymin=.lower, ymax=.upper, fill=Region, group=interaction(Region, hilo)), alpha=0.2) +
  geom_line(aes(color=Region, group=interaction(Region, hilo))) + theme_test() + facet_wrap(~pair)  +
  ylab("expected correlation") + 
  labs(caption="Fig. Specific expected synchrony response to mean SPEI.") +
  theme(plot.caption = element_text(hjust = 0))

clim_summ <- expand_grid(Region=unique(summary_t$Region), hilo=unique(summary_t$hilo), pair=unique(summary_t$pair), spei12_mean=quantile(summary_t$spei12_mean, probs=c( .5)), spei12_sd=quantile(summary_t$spei12_sd, probs=c(.025, .25, .5, .75, .975))) %>% 
  mutate(Site=paste0(Region,hilo)) %>% 
  filter(Site!="Mammothlow") %>% 
  add_fitted_draws(clim_mod, re_formula = NULL, allow_new_levels=TRUE) %>% 
  median_qi()

ggplot(clim_summ, aes(spei12_sd, .value)) +
  geom_ribbon(aes(ymin=.lower, ymax=.upper, fill=Region, group=interaction(Region, hilo)), alpha=0.2) +
  geom_line(aes(color=Region, group=interaction(Region, hilo))) + theme_test() + facet_wrap(~pair)  +
  ylab("expected correlation") + 
  labs(caption="Fig. Specific expected synchrony response to SPEI variablity.") +
  theme(plot.caption = element_text(hjust = 0))



```

```{r misc}

## should go with changes through time model
clim_summ <- expand_grid(Region=unique(summary_t$Region), hilo=unique(summary_t$hilo), pair=unique(summary_t$pair), spei12_mean=quantile(summary_t$spei12_mean, probs=c( .5)), spei12_sd=quantile(summary_t$spei12_sd, probs=c(.5))) %>%
  mutate(Site=paste0(Region,hilo)) %>%
  filter(Site!="Mammothlow") %>%
  add_fitted_draws(clim_mod, re_formula = NULL, allow_new_levels=TRUE) %>%
  compare_levels(.value, by=pair) %>% 
  median_qi(.width=c(0.5, 0.95)) %>% 
  mutate(pair1=str_sub(pair, 1,5),
         pair1_spp1=str_sub(pair, 1,2),
         pair1_spp2=str_sub(pair, 4,5),
         pair1_comp=ifelse(pair1_spp1==pair1_spp2, "intra", "inter"),
         pair2=str_sub(pair, 9,13),
         pair2_spp1=str_sub(pair, 9,10),
         pair2_spp2=str_sub(pair, 12,13),
         pair2_comp=ifelse(pair2_spp1==pair2_spp2, "intra", "inter"),
         valid=ifelse(pair1_comp=="intra", str_detect(pair2, pair1_spp1), str_detect(pair1, pair2_spp1))) %>% 
  filter(pair1_comp!=pair2_comp, valid==TRUE) %>% 
  mutate(new_pair=ifelse(pair1_comp=="intra", str_c(pair1, pair2, sep=" - "), str_c(pair2, pair1, sep=" - ")),
         .value=ifelse(pair1_comp=="intra", .value, .value*-1),
         .upper=ifelse(pair1_comp=="intra", .upper, .upper*-1),
         .lower=ifelse(pair1_comp=="intra", .lower, .lower*-1))

ggplot(clim_summ, aes(x=.value, y=new_pair, color=Region, group=interaction(Region, hilo))) +
  geom_pointintervalh(position = position_dodgev(height=0.5)) + theme_test() +
  geom_vline(xintercept = 0, linetype="dashed")  +
  xlab("expected correlation")

ggplot(clim_summ, aes(x=.value, y=interaction(hilo, Region), color=new_pair)) +
  geom_pointintervalh(position = position_dodgev(height=0.5)) + theme_test() +
  geom_vline(xintercept = 0, linetype="dashed")  +
  xlab("expected correlation")
```



