---
title: "Synchrony from multivariate models"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
author: "Teresa Bohner"
date: "3/30/2020"
output: 
  html_document:
    toc: true
    collapsed: false
    toc_float: true
    toc_depth: 4
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
```

```{r packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(brms)
library(tidybayes)
library(viridis)
library(scales)
library(cowplot)
library(ggridges)
```

## Overview

### Questions
- How does synchrony of tree growth vary across regional/elevation gradients?
-- Prediction: higher synchrony in lower latitude and lower elevation (drier) populations (SIASH, dendrochronological principles)

- Within populations is intraspecific synchrony greater than interspecific synchrony? Between specific species pairs? With increasing distance?
-- Prediction: higher synchrony among intraspecific pairs, potentially higher synchrony between pine pairs versus pine-fir pairs (successional stages), and lower synchrony with increasing distance

- How has synchrony changed through time? Are changes more dramatic in certain populations? Is increased or decreased synchrony associated with certain environmental variables?
-- Prediction: synchrony has increased with time, increased synchrony associated with drier, more variable time windows, changes are more pronounced in xeric populations.

Synchrony is explored here through residual correlations betweeen trees from multivariate models of tree growth.
```{r data}
precip <- read.csv("Processed Data/precip_temp_spei.csv")

timeseq <- seq(1910, 2010, by=10)

precip_summary <- expand_grid(precip, timeseq) %>% 
  filter(Year-timeseq>-11, Year-timeseq<11) %>% 
  group_by(Site, siteno, Neighborhood, hilo, timeseq) %>% 
  summarise_at(vars(total_ppt_mm, mean_temp_C, spei12), list(mean=mean, sd=sd))

rings_field <- read.csv("Processed Data/matched_rings_field.csv")  %>% 
  left_join(precip) %>% 
  group_by(Site, Neighborhood, Species) %>% 
  mutate(treeID=as.numeric(droplevels(as.factor(tree.uniqueID)))) %>% ## counter id if we want it rather than unique ID
  arrange(Site, Neighborhood, Species, ID) 
  

region <- rings_field %>% 
  ungroup() %>% 
  dplyr::select(c(Site, Region, hilo)) %>% 
  group_by(Site) %>% 
  summarize_all(first)

ids <- ungroup(rings_field) %>% 
  dplyr::select(c(Site, Neighborhood, tree.uniqueID)) %>% 
  group_by(Site, Neighborhood, tree.uniqueID) %>%  summarize_all(first)

bigvars2 <- read.csv("Processed Data/synchrony_multivar.csv")
bigdraws2 <- read.csv("Processed Data/synchrony_multivar_draws.csv")
bigvars_t2 <- read.csv("Processed Data/synchrony_decadal.csv") 
  
```

## Q2 Intra-interspecific competition


``` {r funs_and_manip}
## function to extract data at the species level----
spp_subsetter <- function(data, sp) {
  a <- filter(data, pair==paste(sort(c(sp, "ac")), collapse="-")) %>%
    mutate(focal_sp=sp,
           comp_sp="ac")
  b <- filter(data, pair==paste(sort(c(sp, "pj")), collapse="-")) %>%
    mutate(focal_sp=sp,
           comp_sp="pj")
  c <- filter(data, pair==paste(sort(c(sp, "pl")), collapse="-")) %>%
    mutate(focal_sp=sp,
           comp_sp="pl")
  d <- filter(data, pair==paste(sort(c(sp, "pp")), collapse="-")) %>%
    mutate(focal_sp=sp,
           comp_sp="pp")
  
  bind_rows(a,b,c,d) %>%
    mutate(comp_sp=ifelse(comp_sp==focal_sp, "self", comp_sp))
}

## summary data for median correlation
summary <- group_by(bigvars2, spp1, spp2, comp) %>% 
  summarise(med=median(.value))

## subset species----
acdat <- spp_subsetter(bigvars2, "ac")
pjdat <- spp_subsetter(bigvars2, "pj")
pldat <- spp_subsetter(bigvars2, "pl")
ppdat <- spp_subsetter(bigvars2, "pp")


acdat <- spp_subsetter(bigdraws2, "ac")
pjdat <- spp_subsetter(bigdraws2, "pj")
pldat <- spp_subsetter(bigdraws2, "pl")
ppdat <- spp_subsetter(bigdraws2, "pp")

```

*note difference in distribution with when keeping all of the draws versus summarizing median correlations. 

```{r fig}

## matrix style figure
ggplot(bigvars2, aes(.value, color=comp, fill=comp)) +
  geom_density(position = "identity", alpha=0.4) +
  theme_test() +
  geom_vline(data=summary, aes(xintercept = med), linetype="dashed") +
  facet_grid(spp2~spp1)

ggplot(bigdraws2, aes(.value, color=comp, fill=comp)) +
  geom_density(position = "identity", alpha=0.4) +
  theme_test() +
  geom_vline(data=summary, aes(xintercept = med), linetype="dashed") +
  facet_grid(spp2~spp1)

# show_col(viridis_pal()(5)) # get the hex codes for the colors

cols <- c("self"="#440154FF", "pj"="#3B528BFF", "pl"="#21908CFF", "pp"="#5DC863FF", "ac"="#FDE725FF") ## keep species cols consistent

p1 <- ggplot(acdat, aes(.value, color=comp_sp, fill=comp_sp)) +
  geom_density(position = "identity", alpha=0.5) +
  theme_test() +
  ggtitle("A. concolor") +
  scale_fill_manual(values=cols) +
  scale_color_manual(values=cols) +
  theme(legend.position = 'none') +
  xlab("Correlation")

# acdat2 <- left_join(acdat, region) 
# 
# ggplot(acdat2, aes(.value, color=comp_sp, fill=comp_sp)) +
#   geom_density(position = "identity", alpha=0.5) +
#   theme_test() +
#   ggtitle("A. concolor") +
#   scale_fill_manual(values=cols) +
#   scale_color_manual(values=cols) +
#   theme(legend.position = 'none') +
#   facet_grid(Region~hilo)+
#   xlab("Correlation")

p2 <- ggplot(pjdat, aes(.value, color=comp_sp, fill=comp_sp)) +
  geom_density(position = "identity", alpha=0.5) +
  theme_test() +
  ggtitle("P. jeffreyi") +
  scale_fill_manual(values=cols) +
  scale_color_manual(values=cols) +
  theme(legend.position = 'none') +
  xlab("Correlation")

p3 <- ggplot(pldat, aes(.value, color=comp_sp, fill=comp_sp)) +
  geom_density(position = "identity", alpha=0.5) +
  theme_test() +
  ggtitle("P. lambertiana") +
  scale_fill_manual(values=cols) +
  scale_color_manual(values=cols) +
  theme(legend.position = 'none') +
  xlab("Correlation")

p4 <- ggplot(ppdat, aes(.value, color=comp_sp, fill=comp_sp)) +
  geom_density(position = "identity", alpha=0.5) +
  theme_test() +
  ggtitle("P. ponderosa") +
  scale_fill_manual(values=cols) +
  scale_color_manual(values=cols) +
  theme(legend.position = 'none') +
  xlab("Correlation")

dummy <- ggplot(bind_rows(acdat, pjdat), aes(.value, color=comp_sp, fill=comp_sp)) +
  geom_density(alpha=0.5) +
  theme_test() +
  ggtitle("P. lambertiana") +
  scale_fill_manual(values=cols) +
  scale_color_manual(values=cols) +
  labs(fill = "Competitor", color="Competitor")

legend <- cowplot::get_legend(dummy)
# plot_grid(legend)

plot_grid(plot_grid(p1,p2,p3,p4, ncol=2), legend, ncol=2, rel_widths = c(1,.2))
```

## Q3 Synchrony through time
Used 20 year time blocks overlapping 10 years



```{r time_manip}
## subset species----
acdat <- spp_subsetter(bigvars_t2, "ac")
pjdat <- spp_subsetter(bigvars_t2, "pj")
pldat <- spp_subsetter(bigvars_t2, "pl")
ppdat <- spp_subsetter(bigvars_t2, "pp")

cols <- c("self"="#440154FF", "pj"="#3B528BFF", "pl"="#21908CFF", "pp"="#5DC863FF", "ac"="#FDE725FF") ## keep species cols consistent

ggplot(acdat, aes(x=.value, y=as.factor(time), fill=comp_sp)) +
  stat_density_ridges(quantile_lines = TRUE, quantiles = 2,  linetype="dashed") +
  stat_density_ridges(alpha=0) +
  facet_grid(~comp_sp) +
  theme_test() +
  ggtitle("A. concolor") +
  scale_fill_manual(values=cols) +
  theme(legend.position = 'none') +
  xlab("Correlation") +
  ylab("20 year block (year +/- 10)")

ggplot(pjdat, aes(x=.value, y=as.factor(time), fill=comp_sp)) +
  stat_density_ridges(quantile_lines = TRUE, quantiles = 2,  linetype="dashed") +
  stat_density_ridges(alpha=0) +
  facet_grid(~comp_sp) +
  theme_test() +
  ggtitle("P. jeffreyi") +
  scale_fill_manual(values=cols) +
  theme(legend.position = 'none') +
  xlab("Correlation") +
  ylab("20 year block (year +/- 10)")

ggplot(pldat, aes(x=.value, y=as.factor(time), fill=comp_sp)) +
  stat_density_ridges(quantile_lines = TRUE, quantiles = 2,  linetype="dashed") +
  stat_density_ridges(alpha=0) +
  facet_grid(~comp_sp) +
  theme_test() +
  ggtitle("P. lambertiana") +
  scale_fill_manual(values=cols) +
  theme(legend.position = 'none') +
  xlab("Correlation") +
  ylab("20 year block (year +/- 10)")

ggplot(ppdat, aes(x=.value, y=as.factor(time), fill=comp_sp)) +
  stat_density_ridges(quantile_lines = TRUE, quantiles = 2,  linetype="dashed") +
  stat_density_ridges(alpha=0) +
  facet_grid(~comp_sp) +
  theme_test() +
  ggtitle("P. ponderosa") +
  scale_fill_manual(values=cols) +
  theme(legend.position = 'none') +
  xlab("Correlation") +
  ylab("20 year block (year +/- 10)")

```

```{r climate_time}
# ggplot(precip_summary, aes(timeseq, total_ppt_mm_sd, color=Site)) +
#   geom_point() 
# 
# ggplot(precip_summary, aes(timeseq, total_ppt_mm_mean, color=Site)) +
#   geom_point()

ggplot(precip_summary, aes(total_ppt_mm_mean, total_ppt_mm_sd, color=timeseq)) +
  geom_point() +
  theme_test() 

# 
# ggplot(precip_summary, aes(timeseq, mean_temp_C_sd, color=Site)) +
#   geom_point()
# 
# ggplot(precip_summary, aes(timeseq, mean_temp_C_mean, color=Site)) +
#   geom_point()

ggplot(precip_summary, aes(mean_temp_C_mean, mean_temp_C_sd, color=timeseq)) +
  geom_point()+
  theme_test() 

# ggplot(precip_summary, aes(timeseq, spei12_sd, color=Site)) +
#   geom_point()
# 
# ggplot(precip_summary, aes(timeseq, spei12_mean, color=Site)) +
#   geom_point()

ggplot(precip_summary, aes(spei12_mean, spei12_sd, color=timeseq)) +
  geom_point()+
  theme_test() 


```



```{r time_models}
alldata <- bigvars_t2 %>% 
  select(.value, tree1, tree2, spp1, spp2, pair, comp, Site, Neighborhood, time) %>% 
  rename(corr=.value) %>% 
  left_join(precip_summary, by=c("Site","Neighborhood", "time"="timeseq"))

# mod <- brm(corr~time + (time|pair), data=alldata, cores=4)

```